---
title: "Supplementary Material"
output:
  pdf_document:
    toc: true
    fig_caption: true
    latex_engine: lualatex
    number_sections: false
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \floatplacement{table}{H}
  - \floatplacement{verbatim}{H}
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  - \usepackage{lscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
---

```{r echo = FALSE}

knitr::opts_chunk$set(echo = FALSE, fig.width = 11, fig.height = 9, warning = FALSE, message = FALSE, fig.pos = 'H')
library("kableExtra")
library("caper")
library("tidyverse")
library("cowplot")
library("RColorBrewer")
library("ape")
library("RColorBrewer")
library("geiger")
library("phytools")
library("ggrepel")
library("patchwork")
library("plyr")
library("here")
here::i_am("manuscript/new_supp_mat.Rmd")
rm(list = ls())
load(here::here("output/main_results.RData"))
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}
options(digits = 3)

```

\beginsupplement

\blandscape

# Appendix S1 - Comparison with background rates

```{r fig.height = 6, fig.cap = "Relationship between number of species and age for each lineage when compared to the confidence intervals based on the global diversification rate of seed plants. The solid and dashed lines represent the expected richness for $\\epsilon$ = 0 and $\\epsilon$ = 0.9, respectively. The color of the points represents the mycorrhizal state for the 60% threshold."}

## Importing tree
fulltree <- read.tree(here::here("data/fam_tree_family_full.tre"))
fulltree$node.label <- NULL
fulltree.vasc <- read.tree(here::here("data/Vascular_Plants_rooted.dated.tre"))

## Importing results from random datasets
fullresults <- read.csv(here::here("output/fit_data_random_datasets.csv"))

## Importing table with rates for plotting
family.data.gen <- read.csv(here::here("output/simulated_datasets/random_data_09986.csv"), stringsAsFactors = FALSE)
family.data.gen$family[family.data.gen$family == "Leguminosae"] <- "Fabaceae"
family.data.gen$family[family.data.gen$family == "Compositae"] <- "Asteraceae"

age.data <- read.csv(here::here("data/data_all_families.csv"), sep = ";")

## Removing families with unknown mycorrhizal type
family.data.gen <- family.data.gen[-which(family.data.gen$UNK.perc == 1),]

family.data.gen$stem.age <- age.data$stem.age[match(family.data.gen$family, age.data$familia)]
family.data.gen$r.e0 <- bd.ms(time = family.data.gen$stem.age, n = family.data.gen$rich, crown = FALSE, epsilon = 0)
family.data.gen$r.e05 <- bd.ms(time = family.data.gen$stem.age, n = family.data.gen$rich, crown = FALSE, epsilon = 0.5)
family.data.gen$r.e09 <- bd.ms(time = family.data.gen$stem.age, n = family.data.gen$rich, crown = FALSE, epsilon = 0.9)
family.data.gen$shannon <- vegan::diversity(family.data.gen[, 3:7])

## Calculating expected limits from all vascular plants
stem.age.vasc <- max(branching.times(fulltree.vasc)) - findMRCA(fulltree.vasc, c(fulltree.vasc$tip.label[621:622]), type = "height")

r.vasc.stem.r0 <- bd.ms(time = stem.age.vasc, n = sum(family.data.gen$rich), crown = FALSE, epsilon = 0)
r.vasc.stem.r05 <- bd.ms(time = stem.age.vasc, n = sum(family.data.gen$rich), crown = FALSE, epsilon = 0.5)
r.vasc.stem.r0.9 <- bd.ms(time = stem.age.vasc, n = sum(family.data.gen$rich), crown = FALSE, epsilon = 0.9)

limits.vasc.stem <- data.frame(
    time = seq(1, 300, by = 1),
    t(mapply(stem.limits, seq(1, 300, by = 1), r = r.vasc.stem.r0, epsilon = 0)),
    t(mapply(stem.limits, seq(1, 300, by = 1), r = r.vasc.stem.r05, epsilon = 0.5)),
    t(mapply(stem.limits, seq(1, 300, by = 1), r = r.vasc.stem.r0.9, epsilon = 0.9))
    )
names(limits.vasc.stem) <- c("age", "lb.0", "ub.0", "lb.05", "ub.05", "lb.09", "ub.09")

## Stem age
ggplot(data = limits.vasc.stem) +
    geom_line(aes(x = age, y = lb.0)) +
    geom_line(aes(x = age, y = ub.0)) +
    geom_line(aes(x = age, y = lb.09), linetype = "dashed") +
    geom_line(aes(x = age, y = ub.09), linetype = "dashed") +
    geom_point(data = family.data.gen, aes(x = stem.age, y = rich, colour = type.60), size = 2.5) +
    #geom_label_repel(data = family.data.gen, aes(x = stem.age, y = rich, label = family, colour = type.60), segment.alpha = 0.5) +
    #xlim(-50, 300) +
    scale_y_log10(breaks = c(10, 1000, 10000)) +
    labs(x = "Age of Clade (MY)", y = "Number of Species") +
    theme_cowplot() +
    theme(legend.position = "bottom") +
    scale_colour_brewer(palette = "Dark2") +
    labs(colour = "Mycorrhizal State")

```

\elandscape

# Appendix S2 - Relationship bewteen mycorrhizal diversity index, species richness and age

```{r fig.cap = "Scatterplots showing the relationship between mycorrhizal diversity index and (a) stem age and (c) species richness. Panels (b) and (d) show the frequency distribution of R\\textsuperscript{2} and p-values of linear models (blue bars) and PGLS (red bars). Dashed lines show the median for each distribution."}

## Age vs Shannon
scatter.age.sh <-
    ggplot(fullresults) +
    geom_abline(mapping = aes(intercept = pgls.int.age.sh, slope = pgls.slope.age.sh, colour = "PGLS"), show.legend = TRUE, alpha = 1) +
    geom_abline(mapping = aes(intercept = lm.int.age.sh, slope = lm.slope.age.sh, colour = "Linear Model"), show.legend = TRUE, alpha = 1) +
    geom_point(data = na.omit(family.data.gen), aes(x = shannon, y = stem.age), size = 2, alpha = 0.5) +
    labs(x = "Mycorrhizal State Diversity Index", y = "Stem Age", col = "Model Type", title = "A") +
    #xlim(0, 1) +
    scale_colour_manual(values = c("PGLS" = brewer.pal(3, "Set1")[1], "Linear Model" = brewer.pal(3, "Set1")[2])) +
    theme_cowplot() +
    theme(legend.position = "bottom")

r2.pgls.age.sh <-
    ggplot(fullresults) +
    geom_histogram(aes(x = pgls.r2.age.sh), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    geom_vline(xintercept = median(fullresults$pgls.r2.age.sh), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    labs(x = bquote('R' ^2), y = "Frequency", title = "B") +
    theme_cowplot()

pvalue.pgls.age.sh <-
    ggplot(fullresults) +
    geom_histogram(aes(x = pgls.pvalue.age.sh), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    geom_vline(xintercept = median(fullresults$pgls.pvalue.age.sh), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value") +
    theme_cowplot() +
    theme(axis.text.x = element_text(size = 6))

r2.lm.age.sh <-
    ggplot(fullresults) +
    geom_histogram(aes(x = lm.r2.age.sh), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    geom_vline(xintercept = median(fullresults$lm.r2.age.sh), colour = brewer.pal(3, "Set1")[2], linetype = "dashed", size = 1.5) +
    labs(y = element_blank(), x = bquote('R' ^2)) +
    theme_cowplot()

pvalue.lm.age.sh <-
    ggplot(fullresults) +
    geom_histogram(aes(x = lm.pvalue.age.sh), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    geom_vline(xintercept = median(fullresults$lm.pvalue.age.sh), colour = brewer.pal(3, "Set1")[2], linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value") +
    theme_cowplot()


scatter.rich.sh <-
    ggplot(fullresults) +
    geom_abline(mapping = aes(intercept = pgls.int.rich.sh, slope = pgls.slope.rich.sh), colour = brewer.pal(3, "Set1")[1], show.legend = TRUE, alpha = 1) +
    geom_abline(mapping = aes(intercept = lm.int.rich.sh, slope = lm.slope.rich.sh), colour = brewer.pal(3, "Set1")[2], show.legend = TRUE, alpha = 1) +
    geom_point(data = na.omit(family.data.gen), aes(x = shannon, y = rich), size = 2, alpha = 0.5) +
    labs(x = "Mycorrhizal State Diversity Index", y = "Species Richness per Family", colour = "Model Type", title = "C") +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"), direction = -1) +
    #coord_trans(y = "log10") +
    #scale_y_continuous(breaks = c(10, 50, 100, 500, 1000, 10000, 20000, 30000)) +
    theme_cowplot() +
    theme(legend.position = "top")

r2.pgls.rich.sh <-
    ggplot(fullresults) +
    geom_histogram(aes(x = pgls.r2.rich.sh), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    geom_vline(xintercept = median(fullresults$pgls.r2.rich.sh), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = bquote('R' ^2), title = "D")+
    theme_cowplot()

pvalue.pgls.rich.sh <-
    ggplot(fullresults) +
    geom_histogram(aes(x = pgls.pvalue.rich.sh), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    geom_vline(xintercept = median(fullresults$pgls.pvalue.rich.sh), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value") +
    theme_cowplot()

r2.lm.rich.sh <-
    ggplot(fullresults) +
    geom_histogram(aes(x = lm.r2.rich.sh), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    geom_vline(xintercept = median(fullresults$lm.r2.rich.sh), colour = brewer.pal(3, "Set1")[2], linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = bquote('R' ^2)) +
    theme_cowplot()

pvalue.lm.rich.sh <-
    ggplot(fullresults) +
    geom_histogram(aes(x = lm.pvalue.rich.sh), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    geom_vline(xintercept = median(fullresults$lm.pvalue.rich.sh), colour = brewer.pal(3, "Set1")[2], linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value") +
    theme_cowplot()

((scatter.age.sh | ((r2.pgls.age.sh + pvalue.pgls.age.sh) / (r2.lm.age.sh + pvalue.lm.age.sh))) + plot_annotation(tag_levels = 'A'))/ ((scatter.rich.sh | ((r2.pgls.rich.sh + pvalue.pgls.rich.sh) / (r2.lm.rich.sh + pvalue.lm.rich.sh))))

```

# Appendix S3 - Randomly adding 20% of misassignment of mycorrhizal type into the data obtained from genus-level list

```{r fig.height = 4, fig.cap = "Distribution of R\\textsuperscript{2}, p-value and slope value distributions for $\\epsilon$ = 0 after adding 20% misassignment of mycorrhizal type. Blue histograms (upper row) indicate results for standard linear models, whereas red histograms (lower row) indicate results for phylogenetic generalized least squares."}

### 20% analysis
rm(list = ls())
load(here::here("output/hist_20perc.RData"))

(hist.lm.pvalue.r0 + hist.lm.r2.r0 + hist.lm.slope.r0) / (hist.pgls.pvalue.r0 + hist.pgls.r2.r0 + hist.pgls.slope.r0)

```

```{r fig.height = 4, fig.cap = "Distribution of R\\textsuperscript{2}, p-value and slope value distributions for $\\epsilon$ = 0.9 after adding 20% misassignment of mycorrhizal type. Blue histograms (upper row) indicate results for standard linear models, whereas red histograms (lower row) indicate results for phylogenetic generalized least squares."}

(hist.lm.pvalue.r09 + hist.lm.r2.r09 + hist.lm.slope.r09) / (hist.pgls.pvalue.r09 + hist.pgls.r2.r09 + hist.pgls.slope.r09)

```

# Appendix S4 - Analyses with the Species-level dataset

```{r}

## This step is mandatory to avoid problems with overlapping object names
rm(list = ls())
load(here::here("output/suppdata_species_noremarks.RData"))
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

## Clean dataset - excluding species with any remark

```{r fig.height = 11, fig.cap = "Relationship between mycorrhizal type and diversification rates using the thresholds (a) 50\\%, (b) 60\\%, (c) 80\\% and (d) 100\\% for MIX state assignment using the species-level dataset after removing species with remarks. Each panel shows diversification rate estimated with $\\epsilon$ = 0 (upper) and diversification rate estimated with $\\epsilon$ = 0.9 (lower). AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates."}

family.data.clean$type.50 <- as.character(family.data.clean$type.50)
family.data.clean$type.60 <- as.character(family.data.clean$type.60)
family.data.clean$type.80 <- as.character(family.data.clean$type.80)
family.data.clean$type.100 <- as.character(family.data.clean$type.100)

box.r0.50 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.50 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))

thresh50 <-
    plot_grid(box.r0.50,
          box.r09.50,
          nrow = 2,
          align = 'v'
          )

box.r0.60 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.60 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh60 <-
    plot_grid(box.r0.60,
          box.r09.60,
          nrow = 2,
          align = 'v'
          )

box.r0.80 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.80 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh80 <-
    plot_grid(box.r0.80,
          box.r09.80,
          nrow = 2,
          align = 'v'
          )

box.r0.100 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.100 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh100 <-
    plot_grid(box.r0.100,
          box.r09.100,
          nrow = 2,
          align = 'v'
          )

plot_grid(thresh50,
          thresh60,
          thresh80,
          thresh100,
          nrow = 2,
          align = 'hv',
          labels = LETTERS[1:4]
          )

```

### Summary statistics using clean species-level dataset

```{r}

phyaov.sum <- data.frame(Threshold = c("50\\%", "60\\%", "80\\%", "100\\%"),
                                   F.r0 = round(c(phyaov.r0.50$F,
                                            phyaov.r0.60$F,
                                            phyaov.r0.80$F,
                                            phyaov.r0.100$F), 3),
                                   pvalue.r0 = round(c(phyaov.r0.50$Pf,
                                                 phyaov.r0.60$Pf,
                                                 phyaov.r0.80$Pf,
                                                 phyaov.r0.100$Pf), 3),
                                   F.r09 = round(c(phyaov.r09.50$F,
                                            phyaov.r09.60$F,
                                            phyaov.r09.80$F,
                                            phyaov.r09.100$F), 3),
                                   pvalue.r09 = round(c(phyaov.r09.50$Pf,
                                                 phyaov.r09.60$Pf,
                                                 phyaov.r09.80$Pf,
                                                 phyaov.r09.100$Pf), 3)
                         )

phyaov.sum[, 2:5] <- apply(phyaov.sum[, 2:5], 2, sig.formatter)
#names(phyaov.sum)[2:5] <- c("F\n(e = 0)", "p-value\n(e = 0)", "F\n(e = 0.9)", "p-value\n(e = 0.9)")

kable(x = phyaov.sum, format = "latex", caption = "Summary statistics for phyANOVA for both values of $\\epsilon$ using the species-level dataset after removing species with remarks to test for significant differences in diversification rates. Significant values are highlighted in bold.", escape = FALSE, digits = 3, col.names = c("Threshold", "F", "p-value", "F", "p-value"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 2, "e = 0.9" = 2))

```

```{r}

std.sum <- data.frame(Threshold = c("50\\%", "60\\%", "80\\%", "100\\%"),
                      F.r0 = round(c(summary(aov.r0.50)[[1]]$F[1],
                               summary(aov.r0.60)[[1]]$F[1],
                               summary(aov.r0.80)[[1]]$F[1],
                               summary(aov.r0.100)[[1]]$F[1]), 3),
                      pvalue.r0 = round(c(summary(aov.r0.50)[[1]]$P[1],
                                    summary(aov.r0.60)[[1]]$P[1],
                                    summary(aov.r0.80)[[1]]$P[1],
                                    summary(aov.r0.100)[[1]]$P[1]), 3),
                      F.r09 = round(c(summary(aov.r09.50)[[1]]$F[1],
                                summary(aov.r09.60)[[1]]$F[1],
                                summary(aov.r09.80)[[1]]$F[1],
                                summary(aov.r09.100)[[1]]$F[1]), 3),
                      pvalue.r09 = round(c(summary(aov.r09.50)[[1]]$P[1],
                                     summary(aov.r09.60)[[1]]$P[1],
                                     summary(aov.r09.80)[[1]]$P[1],
                                     summary(aov.r09.100)[[1]]$P[1]), 3)
                      )

std.sum[, 2:5] <- apply(std.sum[, 2:5], 2, sig.formatter)
#names(std.sum)[2:5] <- c("F (e = 0)", "p-value (e = 0)", "F (e = 0.9)", "p-value (e = 0.9)")

kable(x = std.sum, format = "latex", caption = "Summary statistics for standard ANOVA for both values of $\\epsilon$ using the species-level dataset after removing species with remarks to test for significant differences in diversification rates. Significant values are highlighted in bold.", escape = FALSE, digits = 3, col.names = c("Threshold", "F", "p-value", "F", "p-value"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 2, "e = 0.9" = 2))

```

### Posthoc tests using clean species-level dataset

```{r}

sum.phyaov <- setNames(
    data.frame(
        threshold = rep(c("50\\%", "60\\%", "80\\%", "100\\%"), each = 4),
        Myc.type = rep(c("AM", "EM", "MIX", "NM"), times = 4),
        rbind(phyaov.r0.50$Pt,
              phyaov.r0.60$Pt,
              phyaov.r0.80$Pt,
              phyaov.r0.100$Pt),
        rbind(phyaov.r09.50$Pt,
              phyaov.r09.60$Pt,
              phyaov.r09.80$Pt,
              phyaov.r09.100$Pt)
    ), nm = c("Threshold", "Mycorrhizal.Type", "AM.r0", "EM.r0", "MIX.r0", "NM.r0", "AM.r09", "EM.r09", "MIX.r09", "NM.r09"))
rownames(sum.phyaov) <- NULL

sum.phyaov[, 3:10] <- apply(sum.phyaov[, 3:10], 2, sig.formatter)
#names(sum.phyaov)[3:10] <- c("AM (e = 0)", "EM (e = 0)", "MIX (e = 0)", "NM (e = 0)", "AM (e = 0.9)", "EM (e = 0.9)", "MIX (e = 0.9", "NM (e = 0.9")

kable(x = sum.phyaov, format = "latex", caption = "Pairwise Corrected p-values for phyANOVA using the species-level dataset after removing species with remarks. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3, col.names = c("Threshold", "Mycorrhizal Type", "AM", "EM", "MIX", "NM", "AM", "EM", "MIX", "NM"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" " = 2, "e = 0" = 4, "e = 0.9" = 4))

```

```{r}

sum.aov <- setNames(
    data.frame(
        pairs = names(TukeyHSD(aov.r0.50)$type.50[,4]),
        round(TukeyHSD(aov.r0.50)$type.50[,4], 3),
        round(TukeyHSD(aov.r0.60)$type.60[,4], 3),
        round(TukeyHSD(aov.r0.80)$type.80[,4], 3),
        round(TukeyHSD(aov.r0.100)$type.100[,4], 3),
        round(TukeyHSD(aov.r09.50)$type.50[,4], 3),
        round(TukeyHSD(aov.r09.60)$type.60[,4], 3),
        round(TukeyHSD(aov.r09.80)$type.80[,4], 3),
        round(TukeyHSD(aov.r09.100)$type.100[,4], 3)
        ), nm = c("Types", "50.r0", "60.r0", "80.r0", "100.r0", "50.r09", "60.r09", "80.r09", "100.r09"))

sum.aov[, 2:9] <- apply(sum.aov[, 2:9], 2, sig.formatter)
#names(sum.phyaov)[2:9] <- c("50 (e = 0)", "60 (e = 0)", "80 (e = 0)", "100 (e = 0)", "50 (e = 0.9)", "60 (e = 0.9)", "80 (e = 0.9", "100 (e = 0.9")

kable(x = sum.aov, format = "latex", caption = "Pairwise Corrected p-values for standard ANOVA. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3, col.names = c("Types", "50\\%", "60\\%", "80\\%", "100\\%", "50\\%", "60\\%", "80\\%", "100\\%"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 4, "e = 0.9" = 4))

```

```{r fig.height = 5, fig.cap = "Scatterplots showing the relationship between mycorrhizal diversity index and diversification rates using the species-level dataset after removing species with remarks. Diversification rates were estimated with $\\epsilon$ = 0 (a) and with $\\epsilon$ = 0.9 (b). The red and blue lines indicate the results of a linear model and a phylogenetic generalized least squares (PGLS) fit, respectively."}

raw.r0 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = shannon, y = r.e0.stem), size = 2) +
    geom_abline(data = data.frame(int = coef(lm.r0)[1], sl = coef(lm.r0)[2], col = "a"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    geom_abline(data = data.frame(int = coef(mod.r0)[1], sl = coef(mod.r0)[2], col = "b"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    labs(x = "Mycorrhizae Type Shannon Index", y = "Diversification Rate", col = "Model Type") +
    #xlim(0, 1) +
    theme_cowplot() +
    theme(legend.position = "top") +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"), direction = -1)

raw.r09 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = shannon, y = r.e09.stem), size = 2) +
    geom_abline(data = data.frame(int = coef(lm.r09)[1], sl = coef(lm.r09)[2], col = "a"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    geom_abline(data = data.frame(int = coef(mod.r09)[1], sl = coef(mod.r09)[2], col = "b"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    labs(x = "Mycorrhizae Type Shannon Index", y = "Diversification Rate", col = "Model Type") +
    #xlim(0, 1) +
    theme_cowplot() +
    theme(legend.position = "top") +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"), direction = -1)

plot_grid(raw.r0,
          raw.r09,
          ncol = 2,
          align = 'hv',
          axis = 'tlbr',
          labels = letters[1:4]
          )

```

```{r}

reg.sum <- data.frame(model = c("PGLS", "LM"),
                      pvalue.r0 = c(4.063e-07, 1.676e-07),
                      R2.r0 = c(0.08904, 0.09458),
                      pvalue.r09 = c(3.483e-07, 2.844e-07),
                      R2.r09 = c(0.09007, 0.09109))

reg.sum[, 2:5] <- apply(reg.sum[, 2:5], 2, sig.formatter)

kable(x = reg.sum, format = "latex", caption = "Summary statistics for the phylogenetic and parametric regressions using the species-level dataset after removing species with remarks. Significant values are highlighted in bold.", escape = FALSE, digits = 3, col.names = c("Model", "p-value", "R\\textsuperscript{2}", "p-value", "R\\textsuperscript{2}"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 2, "e = 0.9" = 2))
  
```


## Full species-level dataset - including species with any remark

```{r}

## This step is mandatory to avoid problems with overlapping object names
rm(list = ls())
load(here::here("output/suppdata_species_withremarks.RData"))
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

```{r fig.height = 11, fig.cap = "Relationship between mycorrhizal type and diversification rates using the thresholds (a) 50\\%, (b) 60\\%, (c) 80\\% and (d) 100\\% for MIX state assignment using the species-level dataset with species with remarks. Each panel shows diversification rate estimated with $\\epsilon$ = 0 (upper) and diversification rate estimated with $\\epsilon$ = 0.9 (lower). AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates."}

family.data.clean.rem$type.50 <- as.character(family.data.clean.rem$type.50)
family.data.clean.rem$type.60 <- as.character(family.data.clean.rem$type.60)
family.data.clean.rem$type.80 <- as.character(family.data.clean.rem$type.80)
family.data.clean.rem$type.100 <- as.character(family.data.clean.rem$type.100)

box.r0.rem.50 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.50 != "ER",]) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.rem.50 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.50 != "ER", ]) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))

thresh50 <- 
    plot_grid(box.r0.rem.50,
              box.r09.rem.50,
              nrow = 2,
              align = 'v'
              )

box.r0.rem.60 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.60 != "ER", ]) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.rem.60 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.50 != "ER", ]) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh60 <-
    plot_grid(box.r0.rem.60,
          box.r09.rem.60,
          nrow = 2,
          align = 'v'
          )

box.r0.rem.80 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.80 != "ER", ]) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.rem.80 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.80 != "ER", ]) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))

thresh80 <-
    plot_grid(box.r0.rem.80,
          box.r09.rem.80,
          nrow = 2,
          align = 'v'
          )

box.r0.rem.100 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.100 != "ER", ]) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.rem.100 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.100 != "ER", ]) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))

thresh100 <-
    plot_grid(box.r0.rem.100,
          box.r09.rem.100,
          nrow = 2,
          align = 'v'
          )

plot_grid(thresh50,
          thresh60,
          thresh80,
          thresh100,
          nrow = 2,
          align = 'hv',
          labels = letters[1:4]
          )

```

### Summary statistics using full species-level dataset

```{r}

phyaov.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                         F.r0 = round(c(phyaov.r0.50.rem$F,
                                        phyaov.r0.60.rem$F,
                                        phyaov.r0.80.rem$F,
                                        phyaov.r0.100.rem$F), 3),
                         pvalue.r0 = round(c(phyaov.r0.50.rem$Pf,
                                             phyaov.r0.60.rem$Pf,
                                             phyaov.r0.80.rem$Pf,
                                             phyaov.r0.100.rem$Pf), 3),
                         F.r09 = round(c(phyaov.r09.50.rem$F,
                                         phyaov.r09.60.rem$F,
                                         phyaov.r09.80.rem$F,
                                         phyaov.r09.100.rem$F), 3),
                         pvalue.r09 = round(c(phyaov.r09.50.rem$Pf,
                                              phyaov.r09.60.rem$Pf,
                                              phyaov.r09.80.rem$Pf,
                                              phyaov.r09.100.rem$Pf), 3)
                         )

phyaov.sum[, 2:5] <- apply(phyaov.sum[, 2:5], 2, sig.formatter)
names(phyaov.sum)[2:5] <- c("F (e = 0)", "p-value (e = 0)", "F (e = 0.9)", "p-value (e = 0.9)")

kable(x = phyaov.sum, format = "latex", caption = "Summary statistics for phyANOVA for both values of $\\epsilon$ using the species-level dataset with species with remarks to test for differences in diversification rates. Significant values are highlighted in bold.", escape = FALSE, digits = 3, col.names = c("Threshold", "F", "p-value", "F", "p-value"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 2, "e = 0.9" = 2))

```

```{r}

std.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                      F.r0 = round(c(summary(aov.r0.50.rem)[[1]]$F[1],
                                     summary(aov.r0.60.rem)[[1]]$F[1],
                                     summary(aov.r0.80.rem)[[1]]$F[1],
                                     summary(aov.r0.100.rem)[[1]]$F[1]), 3),
                      pvalue.r0 = round(c(summary(aov.r0.50.rem)[[1]]$P[1],
                                          summary(aov.r0.60.rem)[[1]]$P[1],
                                          summary(aov.r0.80.rem)[[1]]$P[1],
                                          summary(aov.r0.100.rem)[[1]]$P[1]), 3),
                      F.r09 = round(c(summary(aov.r09.50.rem)[[1]]$F[1],
                                      summary(aov.r09.60.rem)[[1]]$F[1],
                                      summary(aov.r09.80.rem)[[1]]$F[1],
                                      summary(aov.r09.100.rem)[[1]]$F[1]), 3),
                      pvalue.r09 = round(c(summary(aov.r09.50.rem)[[1]]$P[1],
                                           summary(aov.r09.60.rem)[[1]]$P[1],
                                           summary(aov.r09.80.rem)[[1]]$P[1],
                                           summary(aov.r09.100.rem)[[1]]$P[1]), 3)
                      )

std.sum[, 2:5] <- apply(std.sum[, 2:5], 2, sig.formatter)
names(std.sum)[2:5] <- c("F (e = 0)", "p-value (e = 0)", "F (e = 0.9)", "p-value (e = 0.9)")

kable(x = std.sum, format = "latex", caption = "Summary statistics for standard ANOVA for both values of $\\epsilon$ using the species-level dataset with species with remarks to test for differences in diversification rates. Significant values are highlighted in bold.", escape = FALSE, digits = 3, col.names = c("Threshold", "F", "p-value", "F", "p-value"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 2, "e = 0.9" = 2))

```

### Posthoc tests using full species-level dataset

```{r}

sum.phyaov <- setNames(
    data.frame(
        threshold = rep(c("50\\%", "60\\%", "80\\%", "100\\%"), each = 4),
        Myc.type = rep(as.character(c("AM", "EM", "MIX", "NM")), times = 4),
        rbind(phyaov.r0.50.rem$Pt,
              phyaov.r0.60.rem$Pt,
              phyaov.r0.80.rem$Pt,
              phyaov.r0.100.rem$Pt),
        rbind(phyaov.r09.50.rem$Pt,
              phyaov.r09.60.rem$Pt,
              phyaov.r09.80.rem$Pt,
              phyaov.r09.100.rem$Pt),
        stringsAsFactors = FALSE
    ), nm = c("Threshold", "Mycorrhizal.Type", "AM.r0", "EM.r0", "MIX.r0", "NM.r0", "AM.r09", "EM.r09", "MIX.r09", "NM.r09"))
rownames(sum.phyaov) <- NULL

sum.phyaov[, 3:10] <- apply(sum.phyaov[, 3:10], 2, sig.formatter)
names(sum.phyaov)[3:10] <- c("AM (e = 0)", "EM (e = 0)", "MIX (e = 0)", "NM (e = 0)", "AM (e = 0.9)", "EM (e = 0.9)", "MIX (e = 0.9)", "NM (e = 0.9)")

kable(x = sum.phyaov, format = "latex", caption = "Pairwise Corrected p-values for phyANOVA using the species-level dataset with species with remarks. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3, col.names = c("Threshold", "Mycorrhizal Type", "AM", "EM", "MIX", "NM", "AM", "EM", "MIX", "NM"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" " = 2, "e = 0" = 4, "e = 0.9" = 4))

```

```{r}

sum.aov <- setNames(
    data.frame(
        pairs = names(TukeyHSD(aov.r0.50.rem)$type.50[,4]),
        round(TukeyHSD(aov.r0.50.rem)$type.50[,4], 3),
        round(TukeyHSD(aov.r0.60.rem)$type.60[,4], 3),
        round(TukeyHSD(aov.r0.80.rem)$type.80[,4], 3),
        round(TukeyHSD(aov.r0.100.rem)$type.100[,4], 3),
        round(TukeyHSD(aov.r09.50.rem)$type.50[,4], 3),
        round(TukeyHSD(aov.r09.60.rem)$type.60[,4], 3),
        round(TukeyHSD(aov.r09.80.rem)$type.80[,4], 3),
        round(TukeyHSD(aov.r09.100.rem)$type.100[,4], 3)
        ), nm = c("Types", "50.r0", "60.r0", "80.r0", "100.r0", "50.r09", "60.r09", "80.r09", "100.r09"))

sum.aov[, 2:9] <- apply(sum.aov[, 2:9], 2, sig.formatter)
#names(sum.aov)[2:9] <- c("50 (e = 0)", "60 (e = 0)", "80 (e = 0)", "100 (e = 0)", "50 (e = 0.9)", "60 (e = 0.9)", "80 (e = 0.9)", "100 (e = 0.9)")

kable(x = sum.aov, format = "latex", caption = "Pairwise Corrected p-values for standard ANOVA. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3, col.names = c("Types", "50\\%", "60\\%", "80\\%", "100\\%", "50\\%", "60\\%", "80\\%", "100\\%"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 4, "e = 0.9" = 4))

```

```{r fig.height = 5, fig.cap = "Scatterplots showing the relationship between mycorrhizal diversity index and diversification rates using the species-level dataset with species with remarks. Diversification rates were estimated with $\\epsilon$ = 0 (a) and with $\\epsilon$ = 0.9 (b). The red and blue lines indicate the results of a linear model and a phylogenetic generalized least squares (PGLS) fit, respectively."}

load(here::here("output/scatter_boxplots_remarks.RData"))

plot_grid(raw.r0.rem,
          raw.r09.rem,
          ncol = 2,
          align = 'hv',
          axis = 'tlbr',
          labels = letters[1:4]
          )

```

```{r}

reg.sum <- data.frame(model = c("PGLS", "LM"),
                      pvalue.r0 = c(5.265e-05, 2.528e-07),
                      R2.r0 = c(0.06444, 0.09354),
                      pvalue.r09 = c(4.013e-05, 4.977e-07),
                      R2.r09 = c(0.06653, 0.08898))

reg.sum[, 2:5] <- apply(reg.sum[, 2:5], 2, sig.formatter)

kable(x = reg.sum, format = "latex", caption = "Summary statistics for the phylogenetic and parametric regressions using the species-level dataset with species with remarks. Significant values are highlighted in bold.", escape = FALSE, digits = 3, col.names = c("Model", "p-value", "R\\textsuperscript{2}", "p-value", "R\\textsuperscript{2}"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 2, "e = 0.9" = 2))

```

# Appendix S5 - Analyses with the genus-level dataset

## Thresholds for mycorrhizal state assignment

```{r}

## This step is mandatory to avoid problems with overlapping object names
rm(list = ls())
load(here::here("output/suppdata_genus_full.RData"))
load(here::here("output/anova_boxplots.RData"))
##source("../R/anova_plots.R")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

\blandscape

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 50\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0. Panel (A): Boxplot with a randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value both phylogenetic (red) and standard (blue) ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r0.50) + (((pvalue.phyanova.r0.50 + pvalue.anova.r0.50) / posthoc.r0.50) + plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 60\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value both phylogenetic (red) and standard (blue) ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r0.60) + (((pvalue.phyanova.r0.60 + pvalue.anova.r0.60) / posthoc.r0.60) + plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 80\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value both phylogenetic (red) and standard (blue) ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r0.80) + (((pvalue.phyanova.r0.80 + pvalue.anova.r0.80) / posthoc.r0.80) + plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 100\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value both phylogenetic (red) and standard (blue) ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r0.100) + (((pvalue.phyanova.r0.100 + pvalue.anova.r0.100) / posthoc.r0.100) + plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 50\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0.9. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value both phylogenetic (red) and standard (blue) ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r09.50) + (((pvalue.phyanova.r09.50 + pvalue.anova.r09.50) / posthoc.r09.50) + plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 60\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0.9. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value both phylogenetic (red) and standard (blue) ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05. This is the same as figure 2 of the main text and was left there only for easy comparison."}

(box.r09.60) + (((pvalue.phyanova.r09.60 + pvalue.anova.r09.60) / posthoc.r09.60) + plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 80\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0.9. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value both phylogenetic (red) and standard (blue) ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r09.80) + (((pvalue.phyanova.r09.80 + pvalue.anova.r09.80) / posthoc.r09.80) + plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 100\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0.9. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value both phylogenetic (red) and standard (blue) ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r09.100) + (((pvalue.phyanova.r09.100 + pvalue.anova.r09.100) / posthoc.r09.100) + plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

\elandscape


### Summary statistics

```{r}

phyaov.sum <- data.frame(Quantile = c("Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"),
                         as.vector(summary(fullresults$phyaov.pvalue.r0.50)),
                         as.vector(summary(fullresults$phyaov.pvalue.r0.60)),
                         as.vector(summary(fullresults$phyaov.pvalue.r0.80)),
                         as.vector(summary(fullresults$phyaov.pvalue.r0.100)),
                         as.vector(summary(fullresults$phyaov.pvalue.r09.50)),
                         as.vector(summary(fullresults$phyaov.pvalue.r09.60)),
                         as.vector(summary(fullresults$phyaov.pvalue.r09.80)),
                         as.vector(summary(fullresults$phyaov.pvalue.r09.100))
                         )

names(phyaov.sum)[2:9] <- c("50\\%", "60\\%", "80\\%", "100\\%", "50\\%", "60\\%", "80\\%", "100\\%")
phyaov.sum[, 2:9] <- apply(phyaov.sum[, 2:9], 2, sig.formatter)
#names(phyaov.sum)[2:5] <- c("F (e = 0)", "p-value (e = 0)", "F (e = 0.9)", "p-value (e = 0.9)")

kable(x = phyaov.sum, format = "latex", caption = "Summary statistics for phyANOVA for both values of $\\epsilon$ using the genus-level dataset including all families to test for differences in diversification rates. Significant values are highlighted in bold.", escape = FALSE, digits = 3, col.names = c("Quantile", "50\\%", "60\\%", "80\\%", "100\\%", "50\\%", "60\\%", "80\\%", "100\\%"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 4, "e = 0.9" = 4))

```

```{r}

aov.sum <- data.frame(Quantile = c("Minimum", "1st Quart", "Median", "Mean", "3rd Quart", "Max"),
                         round(as.vector(summary(fullresults$aov.pvalue.r0.50)), 5),
                         round(as.vector(summary(fullresults$aov.pvalue.r0.60)), 5),
                         round(as.vector(summary(fullresults$aov.pvalue.r0.80)), 5),
                         round(as.vector(summary(fullresults$aov.pvalue.r0.100)), 5),
                         round(as.vector(summary(fullresults$aov.pvalue.r09.50)), 5),
                         round(as.vector(summary(fullresults$aov.pvalue.r09.60)), 5),
                         round(as.vector(summary(fullresults$aov.pvalue.r09.80)), 5),
                         round(as.vector(summary(fullresults$aov.pvalue.r09.100)), 5)
                         )

names(aov.sum)[2:9] <- c("50\\%", "60\\%", "80\\%", "100\\%", "50\\%", "60\\%", "80\\%", "100\\%")
aov.sum[, 2:9] <- apply(aov.sum[, 2:9], 2, sig.formatter)
#names(phyaov.sum)[2:5] <- c("F (e = 0)", "p-value (e = 0)", "F (e = 0.9)", "p-value (e = 0.9)")

kable(x = aov.sum, format = "latex", caption = "Summary statistics for standard ANOVA for both values of $\\epsilon$ using the genus-level dataset including all families to test for differences in diversification rates. Values were rounded to the fifth decimal place, and significant values are highlighted in bold.", escape = FALSE, digits = 3, col.names = c("Quantile", "50\\%", "60\\%", "80\\%", "100\\%", "50\\%", "60\\%", "80\\%", "100\\%"), align = "c") %>%
  kable_styling(##latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 4, "e = 0.9" = 4))

```

### Posthoc tests using genus-level full dataset

\newpage

```{r}

sum.phyaov <- setNames(
    data.frame(
        epsilon = rep(c(0, 0.9), each = 24),
        threshold = rep(rep(c("50\\%", "60\\%", "80\\%", "100\\%"), each = 6), 2),
        state1 <- c("AM", "AM", "EM", "AM", "EM", "MIX"),
        state2 <- c("EM", "MIX", "MIX", "NM", "NM", "NM"),
        plyr::ldply(list(phyaov.r0.50, phyaov.r0.60, phyaov.r0.80, phyaov.r0.100, phyaov.r09.50, phyaov.r09.60, phyaov.r09.80, phyaov.r09.100), function(x){aggregate(na.omit(x)$value, by = list(na.omit(x)$variable, na.omit(x)$type.2), FUN = summary)$x})),
        nm = c("Epsilon", "Threshold", "State 1", "State 2", "Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"))

rownames(sum.phyaov) <- NULL

#sum.phyaov[, 5:10] <- apply(sum.phyaov[, 3:10], 2, sig.formatter)
#names(sum.phyaov)[3:10] <- c("AM (e = 0)", "EM (e = 0)", "MIX (e = 0)", "NM (e = 0)", "AM (e = 0.9)", "EM (e = 0.9)", "MIX (e = 0.9)", "NM (e = 0.9)")

kable(x = sum.phyaov, format = "latex", caption = "Summary statistics for pairwise corrected p-value distributions for phyANOVA using all 10000 genus-level dataset. Values were rounded to the third decimal place.", row.names = FALSE, escape = FALSE, digits = 3, col.names = c("Epsilon", "Threshold", "State 1", "State 2", "Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F)

```

```{r}

sum.aov <- setNames(
    data.frame(
        epsilon = rep(c(0, 0.9), each = 24),
        threshold = rep(rep(c("50\\%", "60\\%", "80\\%", "100\\%"), each = 6), 2),
        state1 <- c("AM", "AM", "EM", "AM", "EM", "MIX"),
        state2 <- c("EM", "MIX", "MIX", "NM", "NM", "NM"),
        plyr::ldply(list(aov.r0.50, aov.r0.60, aov.r0.80, aov.r0.100, aov.r09.50, aov.r09.60, aov.r09.80, aov.r09.100), function(x){aggregate(na.omit(x)$value, by = list(na.omit(x)$variable, na.omit(x)$type.2), FUN = summary)$x})),
        nm = c("Epsilon", "Threshold", "State 1", "State 2", "Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"))

rownames(sum.aov) <- NULL

#sum.aov[, 5:10] <- apply(sum.aov[, 3:10], 2, sig.formatter)
                                        #names(sum.aov)[3:10] <- c("AM (e = 0)", "EM (e = 0)", "MIX (e = 0)", "NM (e = 0)", "AM (e = 0.9)", "EM (e = 0.9)", "MIX (e = 0.9)", "NM (e = 0.9)")

kable(x = sum.aov, format = "latex", caption = "Summary statistics for pairwise corrected p-value distributions for standard ANOVA using all 10000 genus-level dataset. Values were rounded to the third decimal place.", row.names = FALSE, escape = FALSE, digits = 3, col.names = c("Epsilon", "Threshold", "State 1", "State 2", "Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F)

```

```{r fig.height = 5, fig.cap = "Scatterplots showing a randomly selected dataset replicate (out of 10000) to show the relationship between mycorrhizal state diversity index and diversification rates estimated with $\\epsilon$ (relative extinction fraction) = 0.9. The red and blue lines indicate the linear models and phylogenetic generalized least squares (PGLS) fit, respectively, for 1000 datasets of seed plant mycorrhizal states.  Panels b and c show the frequency distribution of R2 and p-values of (b) linear models (blue bars) and (c) PGLS (red bars), dashed lines show the median for each distribution."}

## Scatter - MTDI vs. Net Diversification
load(here::here("output/scatter_r0.RData"))
scatter.mtdi.r0 | ((r2.lm.r0 + pvalue.lm.r0) / (r2.pgls.r0 + pvalue.pgls.r0))

```


```{r}

reg.sum <- data.frame(Quantile = c("Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"),
                      pvalue.r0.phy = as.vector(summary(fullresults$pgls.pvalue.r0)),
                      R2.r0.phy = as.vector(summary(fullresults$pgls.r2.r0)),
                      pvalue.r09.phy = as.vector(summary(fullresults$pgls.pvalue.r09)),
                      R2.r09.phy = as.vector(summary(fullresults$pgls.r2.r09)),
                      pvalue.r0.lm = as.vector(summary(fullresults$lm.pvalue.r0)),
                      R2.r0.lm = as.vector(summary(fullresults$lm.r2.r0)),
                      pvalue.r09.lm = as.vector(summary(fullresults$lm.pvalue.r09)),
                      R2.r09.lm = as.vector(summary(fullresults$lm.r2.r09))
                      )

#reg.sum[, 2:5] <- apply(reg.sum[, 2:5], 2, sig.formatter)

kable(x = reg.sum, format = "latex", caption = "Summary statistics for p-value and R\\textsuperscript{2} for the phylogenetic and parametric regressions using all 10000 genus-level datasets.", escape = FALSE, digits = 3, col.names = c("Quantile", "p-value", "R\\textsuperscript{2}", "p-value", "R\\textsuperscript{2}", "p-value", "R\\textsuperscript{2}", "p-value", "R\\textsuperscript{2}"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
    add_header_above(c(" ", "e = 0" = 2, "e = 0.9" = 2, "e = 0" = 2, "e = 0.9" = 2)) %>%
    add_header_above(c(" ", "PGLS" = 4, "LM" = 4))

```

# Appendix S6 - Replicating main analyses with tree from Harris and Davies (2016)

```{r}

rm(list = ls())
load(here::here("output/harris_davies_results.RData"))
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

```{r fig.height = 7, fig.cap = "Association between placement of families in both Zanne's (left) and Harris and Davies' (right) phylogenies, indicating differences in the topologies."}

cophyloplot(fulltree, tree, assoc = cophylo.myco$assoc, use.edge.length = TRUE, show.tip.label = FALSE, space = 500)

```

```{r fig.height = 5, fig.cap = "Relationship between net diversification calculated in our study and with the rates from Harris and Davies (2016) The rates were estimated by the authors using the same source for species richness as the main analysis, and the stem ages were obtained from their phylogeny. a: $\\epsilon$ = 0; b: $\\epsilon$ = 0.9. Red lines: LM fit; blue dashed lines: identity fit."}

r0.corr <-
    ggplot(data.muj) +
    geom_point(aes(x = new.phylo.r.e0, y = r.e0)) +
    geom_smooth(aes(x = new.phylo.r.e0, y = r.e0), method = "lm", se = FALSE, colour = "red") +
    geom_abline(slope = 1, intercept = 0, colour = "blue", linetype = "dashed") +
    labs(x = "Harris and Davies\n(e = 0)", y = "Zanne et al\n(e = 0)") +
    cowplot::theme_cowplot()

r09.corr <-
    ggplot(data.muj) +
    geom_point(aes(x = new.phylo.r.e09, y = r.e09)) +
    geom_smooth(aes(x = new.phylo.r.e09, y = r.e09), method = "lm", se = FALSE, colour = "red") +
    geom_abline(slope = 1, intercept = 0, colour = "blue", linetype = "dashed") +
    labs(x = "Harris and Davies\n(e = 0.9)", y = "Zanne et al\n(e = 0.9)") +
    cowplot::theme_cowplot()

plot_grid(r0.corr,
          r09.corr,
          ncol = 2,
          align = 'h',
          labels = letters[1:2]
          )

```

```{r fig.cap = "Main analyses replicated using Harris and Davies (2016) phylogeny. The rates were estimated by the authors using the same source for species richness as the main analysis, and the stem ages were obtained from their phylogeny. The models were tested using only the empirical dataset, with no randomization. Panels a and c: $\\epsilon$ = 0; panels b and d: $\\epsilon$ = 0.9. Red lines: PGLS fit; blue dashed lines: LM fit."}

sh.r0.np <-
    ggplot(data = data.muj.plot) +
    geom_point(aes(x = shannon, y = new.phylo.r.e0)) +
    geom_abline(intercept = coef(lm.np.r0)[1], slope = coef(lm.np.r0)[2], colour = brewer.pal(3, "Set1")[2], size = 2) +
    geom_abline(intercept = coef(mod.np.r0)[1], slope = coef(mod.np.r0)[2], colour = brewer.pal(3, "Set1")[1], size = 2) +
    labs(x = "Mycorrhizal Type Diversity Index", y = "NetDiversification\n(e = 0)", colour = "Model") +
    theme_cowplot() +
    theme(legend.position = "top")

sh.r09.np <-
    ggplot(data = data.muj.plot) +
    geom_point(aes(x = shannon, y = new.phylo.r.e09)) +
    geom_abline(intercept = coef(lm.np.r09)[1], slope = coef(lm.np.r09)[2], colour = brewer.pal(3, "Set1")[2], size = 2) +
    geom_abline(intercept = coef(mod.np.r09)[1], slope = coef(mod.np.r09)[2], colour = brewer.pal(3, "Set1")[1], size = 2) +
    labs(x = "Mycorrhizal Type Diversity Index", y = "NetDiversification\n(e = 0.9)", colour = "Model") +
    theme_cowplot() +
    theme(legend.position = "top")


aov.r0.np <-
    ggplot(data = data.aov) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = new.phylo.r.e0, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")))) +
    geom_jitter(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = new.phylo.r.e0, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5) +
    labs(x = "Mycorrhizal Type", y = "NetDiversification\n(e = 0)") +
    scale_colour_manual(values = RColorBrewer::brewer.pal(5, "Set1")[c(1:3, 5)]) +
    theme_cowplot() +
    theme(legend.position = "none")


aov.r09.np <-
    ggplot(data = data.aov) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = new.phylo.r.e09, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")))) +
    geom_jitter(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = new.phylo.r.e09, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5) +
    labs(x = "Mycorrhizal Type", y = "NetDiversification\n(e = 0.9)") +
    scale_colour_manual(values = RColorBrewer::brewer.pal(5, "Set1")[c(1:3, 5)]) +
    theme_cowplot() +
    theme(legend.position = "none")

plot_grid(sh.r0.np,
          sh.r09.np,
          aov.r0.np,
          aov.r09.np,
          ncol = 2,
          align = 'hv',
          labels = letters[1:4]
          )

```

```{r}

np.sum <-
    data.frame(model = c("Phylogenetic", "Linear"),
                     pvalue.r0.reg = c("\\textbf{2.724e-13}", "\\textbf{1.345e-14}"),
                     R2.r0.reg = c(0.1349, 0.1464),
                     pvalue.r0.aov = c("\\textbf{0.0005}", "\\textbf{0.019}"),
                     pvalue.r09.reg = c("\\textbf{3.581e-13}", "\\textbf{1.228e-14}"),
                     R2.r09.reg = c(0.1336, 0.1468),
                     pvalue.r09.aov = c("\\textbf{0.0003}", "\\textbf{0.02}")
                     )

kable(x = np.sum,
      format = "latex", caption = "Summary statistics for linear models and ANOVA using Harris and Davies (2016) phylogeny.", escape = FALSE, digits = 3, col.names = c("Model", "p-value", "R\\textsuperscript{2}", "p-value", "p-value", "R\\textsuperscript{2}", "p-value"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "Regression" = 2, "ANOVA" = 1, "Regression" = 2, "ANOVA" = 1)) %>%
  add_header_above(c(" ", "e = 0" = 3, "e = 0.9" = 3))

```

# Appendix S7 - Phylogenetic signal of diversification rates, age, and richness

```{r}

rm(list = ls())
load(here::here("output/main_results.RData"))
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

```{r fig.cap = "P-value distribution of phylogenetic signal of each mycorrhizal state using different thresholds for classification. Colour indicates each model to which the phylogenetic structure of mycorrhizal type is compared, and vertical dashed line indicates the value of 0.05", fig.height = 10}

## age.data <- read.csv("./data/data_all_families.csv", sep = ";")
## fulltree <- read.tree("./data/fam_tree_family_full.tre")
## fulltree$node.label <- NULL

## nrep = length(list.files("./output/simulated_datasets/"))

## phylosig <- function(x, age, fulltree){
##     print(paste0("Replica ", x, " of ", length(list.files("../output/simulated_datasets/"))))
##     family.data.gen <- read.csv(paste0("../output/simulated_datasets/random_data_", sprintf("%05d", x), ".csv"), stringsAsFactors = FALSE)
##     family.data.gen$family[family.data.gen$family == "Leguminosae"] <- "Fabaceae"
##     family.data.gen$family[family.data.gen$family == "Compositae"] <- "Asteraceae"

##     ## Removing families with unknown mycorrhizal type
##     family.data.gen <- family.data.gen[-which(family.data.gen$UNK.perc == 1),]

##     family.data.gen$stem.age <- age$stem.age[match(family.data.gen$family, age$familia)]
##     family.data.gen$r.e0 <- bd.ms(time = family.data.gen$stem.age, n = family.data.gen$rich, crown = FALSE, epsilon = 0)
##     family.data.gen$r.e09 <- bd.ms(time = family.data.gen$stem.age, n = family.data.gen$rich, crown = FALSE, epsilon = 0.9)

##     family.data.gen$shannon <- vegan::diversity(family.data.gen[, 3:7])

##     family.data.gen <- family.data.gen[-match(c("Orchidaceae", "Ericaceae", "Diapensiaceae"), family.data.gen$family),]

##     tree.pruned <- drop.tip(fulltree, tip = fulltree$tip.label[is.na(match(fulltree$tip.label, family.data.gen$family))])
##     data.pgls <- comparative.data(tree.pruned, family.data.gen, names.col = "family")
    
##     data.phylo.d50 <- family.data.gen[, c("family", "type.50")]
##     data.phylo.d60 <- family.data.gen[, c("family", "type.60")]
##     data.phylo.d80 <- family.data.gen[, c("family", "type.80")]
##     data.phylo.d100 <- family.data.gen[, c("family", "type.100")]

##     ## Threshold 50%

##     data.phylo.d50$dummy.am <- ifelse(data.phylo.d50$type.50 == "AM", 1, 0)
##     data.phylo.d50$dummy.em <- ifelse(data.phylo.d50$type.50 == "EM", 1, 0)
##     data.phylo.d50$dummy.mix <- ifelse(data.phylo.d50$type.50 == "MIX", 1, 0)
##     data.phylo.d50$dummy.nm <- ifelse(data.phylo.d50$type.50 == "NM", 1, 0)

##     d.am.50 <- phylo.d(data.phylo.d50, data.pgls$phy, names.col = family, binvar = dummy.am)
##     d.em.50 <- phylo.d(data.phylo.d50, data.pgls$phy, names.col = family, binvar = dummy.em)
##     d.mix.50 <- phylo.d(data.phylo.d50, data.pgls$phy, names.col = family, binvar = dummy.mix)
##     d.nm.50 <- phylo.d(data.phylo.d50, data.pgls$phy, names.col = family, binvar = dummy.nm)

##     ## Threshold 60%

##     data.phylo.d60$dummy.am <- ifelse(data.phylo.d60$type.60 == "AM", 1, 0)
##     data.phylo.d60$dummy.em <- ifelse(data.phylo.d60$type.60 == "EM", 1, 0)
##     data.phylo.d60$dummy.mix <- ifelse(data.phylo.d60$type.60 == "MIX", 1, 0)
##     data.phylo.d60$dummy.nm <- ifelse(data.phylo.d60$type.60 == "NM", 1, 0)

##     d.am.60 <- phylo.d(data.phylo.d60, data.pgls$phy, names.col = family, binvar = dummy.am)
##     d.em.60 <- phylo.d(data.phylo.d60, data.pgls$phy, names.col = family, binvar = dummy.em)
##     d.mix.60 <- phylo.d(data.phylo.d60, data.pgls$phy, names.col = family, binvar = dummy.mix)
##     d.nm.60 <- phylo.d(data.phylo.d60, data.pgls$phy, names.col = family, binvar = dummy.nm)

##     ## Threshold 80%

##     data.phylo.d80$dummy.am <- ifelse(data.phylo.d80$type.80 == "AM", 1, 0)
##     data.phylo.d80$dummy.em <- ifelse(data.phylo.d80$type.80 == "EM", 1, 0)
##     data.phylo.d80$dummy.mix <- ifelse(data.phylo.d80$type.80 == "MIX", 1, 0)
##     data.phylo.d80$dummy.nm <- ifelse(data.phylo.d80$type.80 == "NM", 1, 0)

##     d.am.80 <- phylo.d(data.phylo.d80, data.pgls$phy, names.col = family, binvar = dummy.am)
##     d.em.80 <- phylo.d(data.phylo.d80, data.pgls$phy, names.col = family, binvar = dummy.em)
##     d.mix.80 <- phylo.d(data.phylo.d80, data.pgls$phy, names.col = family, binvar = dummy.mix)
##     d.nm.80 <- phylo.d(data.phylo.d80, data.pgls$phy, names.col = family, binvar = dummy.nm)

##     ## Threshold 100%

##     data.phylo.d100$dummy.am <- ifelse(data.phylo.d100$type.100 == "AM", 1, 0)
##     data.phylo.d100$dummy.em <- ifelse(data.phylo.d100$type.100 == "EM", 1, 0)
##     data.phylo.d100$dummy.mix <- ifelse(data.phylo.d100$type.100 == "MIX", 1, 0)
##     data.phylo.d100$dummy.nm <- ifelse(data.phylo.d100$type.100 == "NM", 1, 0)

##     d.am.100 <- phylo.d(data.phylo.d100, data.pgls$phy, names.col = family, binvar = dummy.am)
##     d.em.100 <- phylo.d(data.phylo.d100, data.pgls$phy, names.col = family, binvar = dummy.em)
##     d.mix.100 <- phylo.d(data.phylo.d100, data.pgls$phy, names.col = family, binvar = dummy.mix)
##     d.nm.100 <- phylo.d(data.phylo.d100, data.pgls$phy, names.col = family, binvar = dummy.nm)

##     phylod.table <- data.frame(
##         replica = rep(x, 8),
##         Threshold = rep(c("50\\%", "60\\%", "80\\%", "100\\%"), each = 2),
##         model = rep(c("random", "BM"), times = 4),
##         AM = c(unlist(d.am.50[c("Pval1", "Pval0")]),
##                unlist(d.am.60[c("Pval1", "Pval0")]),
##                unlist(d.am.80[c("Pval1", "Pval0")]),
##                unlist(d.am.100[c("Pval1", "Pval0")])),
##         EM = c(unlist(d.em.50[c("Pval1", "Pval0")]),
##                unlist(d.em.60[c("Pval1", "Pval0")]),
##                unlist(d.em.80[c("Pval1", "Pval0")]),
##                unlist(d.em.100[c("Pval1", "Pval0")])),
##         MIX = c(unlist(d.mix.50[c("Pval1", "Pval0")]),
##                 unlist(d.mix.60[c("Pval1", "Pval0")]),
##                 unlist(d.mix.80[c("Pval1", "Pval0")]),
##                 unlist(d.mix.100[c("Pval1", "Pval0")])),
##         NM = c(unlist(d.nm.50[c("Pval1", "Pval0")]),
##                unlist(d.nm.60[c("Pval1", "Pval0")]),
##                unlist(d.nm.80[c("Pval1", "Pval0")]),
##                unlist(d.nm.100[c("Pval1", "Pval0")]))
##     )
##     return(phylod.table)
## }

## doMC::registerDoMC(56)
## phylod.table <- plyr::ldply(1:10000, phylosig, age = age.data, fulltree = fulltree)
load(here::here("output/phylod_table.RData"))
#phylod.table[, 4:7] <- apply(phylod.table[, 4:7], 2, sig.formatter)

phylod.AM <-
    ggplot(data = phylod.table) +
    geom_histogram(aes(x = AM, fill = model), bins = 75, alpha = 0.8) +
    geom_vline(xintercept = 0.05, colour = "darkgrey", linetype = "dashed") +
    facet_grid(factor(gsub("\\", "", Threshold, fixed = TRUE), levels = c("50%", "60%", "80%", "100%")) ~ .) +
    scale_fill_brewer(palette = "Dark2") +
    labs(y = "Frequency", fill = "Model Type") +
    theme_cowplot() +
    theme(legend.position = "bottom")

phylod.EM <-
    ggplot(data = phylod.table) +
    geom_histogram(aes(x = EM, fill = model), bins = 75, alpha = 0.8) +
    geom_vline(xintercept = 0.05, colour = "darkgrey", linetype = "dashed") +
    facet_grid(factor(gsub("\\", "", Threshold, fixed = TRUE), levels = c("50%", "60%", "80%", "100%")) ~ .) +
    scale_fill_brewer(palette = "Dark2") +
    labs(y = element_blank(), fill = "Model Type") +
    theme_cowplot() +
    theme(legend.position = "bottom")

phylod.NM <-
    ggplot(data = phylod.table) +
    geom_histogram(aes(x = NM, fill = model), bins = 75, alpha = 0.8) +
    geom_vline(xintercept = 0.05, colour = "darkgrey", linetype = "dashed") +
    facet_grid(factor(gsub("\\", "", Threshold, fixed = TRUE), levels = c("50%", "60%", "80%", "100%")) ~ .) +
    scale_fill_brewer(palette = "Dark2") +
    labs(y = element_blank(), fill = "Model Type") +
    theme_cowplot() +
    theme(legend.position = "bottom")

phylod.MIX <-
    ggplot(data = phylod.table) +
    geom_histogram(aes(x = MIX, fill = model), bins = 75, alpha = 0.8) +
    geom_vline(xintercept = 0.05, colour = "darkgrey", linetype = "dashed") +
    facet_grid(factor(gsub("\\", "", Threshold, fixed = TRUE), levels = c("50%", "60%", "80%", "100%")) ~ .) +
    scale_fill_brewer(palette = "Dark2") +
    labs(y = element_blank(), fill = "Model Type") +
    theme_cowplot() +
    theme(legend.position = "bottom")

(phylod.AM | phylod.EM | phylod.NM | phylod.MIX) / guide_area() +
    plot_layout(guides = "collect", heights = c(1, 0.05))

## kable_styling(kable(x = phylod.table, format = "latex", caption = "P-values for test of phylogenetic signal (D) of each mycorrhizal type using the genus-level dataset including all families. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

```{r}

kable_styling(kable(x = data.frame(Variable = c("r (epsilon = 0)", "r (epsilon = 0.9)", "Stem Age", "Richness"),
                                   Lambda = c(paste0("\\textbf{", round(summary(phylosig.r0)$param[2], 3), "}"),
                                              paste0("\\textbf{", round(summary(phylosig.r09)$param[2], 3), "}"), 
                                              paste0("\\textbf{", summary(phylosig.age)$param[2], "}"),
                                              summary(phylosig.rich)$param[2]
                                              )),
                    format = "latex", caption = "Phylogenetic signal of the four response variables using the genus-level dataset including all families. Significant values are highlighted in bold.", escape = FALSE), latex_options = "HOLD_position"
              )

```


# Appendix S8 - Percentage of MIX species

```{r fig.width = 11, fig.height = 7, fig.cap = "Distribution of percentage of species in each family that are classified as MIX. Zoom panel indicates the same distribution but excluding families without any MIX species."}

library("ggforce")
mix.sp <- read.csv(here::here("data/family_data_genus_final.csv"), stringsAsFactors = FALSE)
mix.sp$zoom.bin <- "a"
mix.sp$zoom.bin[mix.sp$MIX.raw.perc != 0] <- "b"

fullhist <-
    ggplot(mix.sp) +
    geom_histogram(aes(x = MIX.raw.perc, fill = zoom.bin, colour = zoom.bin), bins = nrow(mix.sp), size = 2) +
    labs(x = "Proportion of MIX species", y = "Frequency", title = "A") +
    scale_colour_manual(values = c("a" = "black", "b" = "red")) +
    scale_fill_manual(values = c("a" = "black", "b" = "red")) +
    theme_cowplot() +
    theme(legend.position = "none")

mix.hist <-
    ggplot(mix.sp[mix.sp$MIX.raw.perc != 0,]) +
    geom_histogram(aes(x = MIX.raw.perc, fill = zoom.bin, colour = zoom.bin), bins = nrow(mix.sp[mix.sp$MIX.raw.perc != 0,]), colour = "red", fill = "red") +
    labs(x = "Proportion of MIX species", y = "Frequency", title = "B") +
    theme_cowplot()

fullhist + annotation_custom(grob = ggplotGrob(mix.hist), xmin = 0.3, xmax = 1, ymin = 100, ymax = 300)

## ggsave(filename = "./output/figs/hist_mix_proportion.pdf")

```

# Replicating main analyses with all trees from Ramirez-Barahona \textit{et al.} (2020)

```{r rambarah_results, echo = FALSE, fig.cap = "Main analyses replicated using Ramirez-Barahona \textit{et al.} (2020) phylogeny. The rates were estimated by the authors using the same source for species richness as the main analysis, and the stem ages were obtained from their phylogeny. Panels A: scatterplot with linear and phylogenetic regressions plotted; panel B: histograms of R2 and p-value of linear and phylogenetic regressions. Red lines/bars: PGLS fit; blue lines/bars: LM fit."}

age.data <- read.csv(here::here("data/data_all_families.csv"), sep = ";")
RB.tree.RC.complete <- read.nexus(here::here("data/ramirez_barahona_data/ramirez_barahona_RC_complete_MCCv_2.tre"))
RB.tree.RC.conservative <- read.nexus(here::here("data/ramirez_barahona_data/ramirez_barahona_RC_conservative_MCCv_2.tre"))
RB.tree.UC.complete <- read.nexus(here::here("data/ramirez_barahona_data/ramirez_barahona_UC_complete_MCCv_2.tre"))
RB.tree.UC.conservative <- read.nexus(here::here("data/ramirez_barahona_data/ramirez_barahona_UC_conservative_MCCv_2.tre"))
RB.tree.CC.complete <- read.nexus(here::here("data/ramirez_barahona_data/ramirez_barahona_CC_complete_MCCv_2.tre"))
RB.tree.CC.conservative <- read.nexus(here::here("data/ramirez_barahona_data/ramirez_barahona_CC_conservative_MCCv_2.tre"))

## Extracting families from tip information

RB.tip.list <- RB.tree.CC.complete$tip.label
RB.family.list <- unique(setNames(sapply(RB.tip.list, function(x){strsplit(x, split = "_")[[1]][2]}), NULL))

## Replacing species name by the family, and dropping repeated tips
RB.tree.CC.complete.pruned <- drop.tip(RB.tree.CC.complete, tip = grep(RB.family.list[1], RB.tree.CC.complete$tip.label)[-1])
for(i in 2:length(RB.family.list)){
    RB.tree.CC.complete.pruned <- drop.tip(RB.tree.CC.complete.pruned, tip = grep(RB.family.list[i], RB.tree.CC.complete.pruned$tip.label)[-1])
}
RB.tree.CC.complete.pruned$tip.label <- setNames(sapply(RB.tree.CC.complete.pruned$tip.label, function(x){strsplit(x, split = "_")[[1]][2]}), NULL)

rb.CC.complete.ages <- read.csv(here::here("data/ramirez_barahona_data/ramirez_barahona_Ages_CC_complete.csv"), as.is = TRUE)

## Missing families from Ramrez-Barahona \textit{et al.} 2020
age.data$familia[is.na(match(age.data$familia, rb.CC.complete.ages$Family))]

## Adding ages from Ramrez-Barahona \textit{et al.} 2020
age.data$rb.CC.complete.stem <- rb.CC.complete.ages$Stem_BEAST[match(age.data$familia, rb.CC.complete.ages$Family)]

## Importing results from random datasets
load(here::here("output/RB_results_RC_complete.RData"))
load(here::here("output/RB_results_RC_conservative.RData"))
load(here::here("output/RB_results_UC_complete.RData"))
load(here::here("output/RB_results_UC_conservative.RData"))
load(here::here("output/RB_results_CC_complete.RData"))
load(here::here("output/RB_results_CC_conservative.RData"))
nrep <- 10000

fullresults.RC.complete <- ldply(1:nrep, function(x){results.RC.complete[[x]][[1]]})
fullresults.RC.conservative <- ldply(1:nrep, function(x){results.RC.conservative[[x]][[1]]})
fullresults.UC.complete <- ldply(1:nrep, function(x){results.UC.complete[[x]][[1]]})
fullresults.UC.conservative <- ldply(1:nrep, function(x){results.UC.conservative[[x]][[1]]})
fullresults.CC.complete <- ldply(1:nrep, function(x){results.CC.complete[[x]][[1]]})
fullresults.CC.conservative <- ldply(1:nrep, function(x){results.CC.conservative[[x]][[1]]})

fullresults <- bind_rows(list(fullresults.RC.complete,
                              fullresults.RC.conservative,
                              fullresults.UC.complete,
                              fullresults.UC.conservative,
                              fullresults.CC.complete,
                              fullresults.CC.conservative))

fullresults$tree <- rep(c("RC Complete", "RC Conservative", "UC Complete", "UC Conservative", "CC Complete", "CC Conservative"), each = 10000)

## Scatter - MTDI vs. Net Diversification

## Using replica 9986 (with R2 closer to the median) just to represent points

scatter.mtdi.r09.rambarah <-
    ggplot(fullresults) +
    geom_abline(mapping = aes(intercept = pgls.int.r09, slope = pgls.slope.r09, colour = brewer.pal(3, "Set1")[1]), show.legend = TRUE, alpha = 1) +
    geom_abline(mapping = aes(intercept = lm.int.r09, slope = lm.slope.r09, colour = brewer.pal(3, "Set1")[2]), show.legend = TRUE, alpha = 1) +
    geom_point(data = na.omit(family.data.gen), aes(x = shannon, y = r.e09), size = 2, alpha = 0.5) +
    labs(x = "Mycorrhizal State Diversity Index", y = "Diversification Rate", col = "Model Type", title = "A") +
    facet_wrap(. ~ tree, ncol = 2) +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"), direction = -1) +
    theme_cowplot() +
    theme(legend.position = "top")

r2.r09.rambarah <-
    ggplot(fullresults) +
    geom_histogram(aes(x = lm.r2.r09), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    geom_histogram(aes(x = pgls.r2.r09), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    ## geom_vline(aes(xintercept = median(lm.r2.r09)), colour = brewer.pal(3, "Set1")[2], linetype = "dashed", size = 1.5) +
    ## geom_vline(aes(xintercept = median(pgls.r2.r09)), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    facet_wrap(. ~ tree, ncol = 2) +
    labs(y = "Frequency", x = bquote('R' ^2), title = "B") +
    theme_cowplot()

pvalue.r09.rambarah <-
    ggplot(fullresults) +
    geom_histogram(aes(x = log10(lm.pvalue.r09)), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    geom_histogram(aes(x = log10(pgls.pvalue.r09)), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    ## geom_vline(xintercept = log10(median(fullresults$lm.pvalue.r09)), colour = brewer.pal(3, "Set1")[2], linetype = "dashed", size = 1.5) +
    ## geom_vline(xintercept = log10(median(fullresults$pgls.pvalue.r09)), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value (log10)") +
    facet_wrap(. ~ tree, ncol = 2) +
    theme_cowplot()

(scatter.mtdi.r09.rambarah | (r2.r09.rambarah + pvalue.r09.rambarah)) + plot_annotation(tag_levels = "A")

```

```{r echo = FALSE, fig.cap = "Analysis of the relationship between age and MTDI replicated using Ramirez-Barahona \textit{et al.} (2020) phylogeny. The rates were estimated by the authors using the same source for species richness as the main analysis, and the stem ages were obtained from their phylogeny. Panel A: scatterplots with linear and phylogenetic regressions plotted; panel B: histograms of R2 and p-value of linear and phylogenetic regressions. Red lines/bars: PGLS fit; blue lines/bars: LM fit."}

## Age vs Shannon
scatter.age.sh.rambarah <-
    ggplot(fullresults) +
    geom_abline(mapping = aes(intercept = pgls.int.age.sh, slope = pgls.slope.age.sh), colour = brewer.pal(3, "Set1")[1], show.legend = TRUE, alpha = 0.01) +
    geom_abline(mapping = aes(intercept = lm.int.age.sh, slope = lm.slope.age.sh), colour = brewer.pal(3, "Set1")[2], show.legend = TRUE, alpha = 0.01) +
    geom_point(data = na.omit(family.data.gen), aes(x = shannon, y = stem.age), size = 2, alpha = 0.5) +
    labs(x = "Mycorrhizal State Shannon Index", y = "Stem Age", col = "Model Type") +
    facet_wrap(. ~ tree, ncol = 2) +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"), direction = -1) +
    theme_cowplot() +
    theme(legend.position = "top")

r2.age.sh.rambarah <-
    ggplot(fullresults) +
    geom_histogram(aes(x = pgls.r2.age.sh), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    geom_histogram(aes(x = lm.r2.age.sh), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    ## geom_vline(xintercept = median(fullresults$pgls.r2.age.sh), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    ## geom_vline(xintercept = median(fullresults$lm.r2.age.sh), colour = brewer.pal(3, "Set1")[2], linetype = "dashed", size = 1.5) +
    labs(x = bquote('R' ^2), y = "Frequency") +
    facet_wrap(. ~ tree, ncol = 2) +
    theme_cowplot()

pvalue.age.sh.rambarah <-
    ggplot(fullresults) +
    geom_histogram(aes(x = pgls.pvalue.age.sh), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    geom_histogram(aes(x = lm.pvalue.age.sh), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    ## geom_vline(xintercept = median(fullresults$pgls.pvalue.age.sh), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    ## geom_vline(xintercept = median(fullresults$lm.pvalue.age.sh), colour = brewer.pal(3, "Set1")[2], linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value") +
    facet_wrap(. ~ tree, ncol = 2) +
    theme_cowplot()

(scatter.age.sh.rambarah | (r2.age.sh.rambarah + pvalue.age.sh.rambarah)) + plot_annotation(tag_levels = "A")

```

```{r echo = FALSE, fig.cap = "Analysis of the relationship between richness and MTDI replicated using Ramirez-Barahona \textit{et al.} (2020) phylogeny. The rates were estimated by the authors using the same source for species richness as the main analysis, and the stem ages were obtained from their phylogeny. Panel A: scatterplots with linear and phylogenetic regressions plotted; panel B: histograms of R2 and p-value of linear and phylogenetic regressions. Red lines/bars: PGLS fit; blue lines/bars: LM fit."}

## Richness vs Shannon
scatter.rich.sh.rambarah <-
    ggplot(fullresults) +
    geom_abline(mapping = aes(intercept = pgls.int.rich.sh, slope = pgls.slope.rich.sh), colour = brewer.pal(3, "Set1")[1], show.legend = TRUE, alpha = 0.01) +
    geom_abline(mapping = aes(intercept = lm.int.rich.sh, slope = lm.slope.rich.sh), colour = brewer.pal(3, "Set1")[2], show.legend = TRUE, alpha = 0.01) +
    geom_point(data = na.omit(family.data.gen), aes(x = shannon, y = rich), size = 2, alpha = 0.5) +
    labs(x = "Mycorrhizal State Shannon Index", y = "Species Richness per Family", col = "Model Type") +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"), direction = -1) +
    facet_wrap(. ~ tree, ncol = 2) +
    #scale_y_continuous(breaks = c(10, 50, 100, 500, 1000, 10000, 20000, 30000)) +
    theme_cowplot() +
    theme(legend.position = "top")

r2.rich.sh.rambarah <-
    ggplot(fullresults) +
    geom_histogram(aes(x = pgls.r2.rich.sh, y = ..density..), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    geom_histogram(aes(x = lm.r2.rich.sh, y = ..density..), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    ## geom_vline(xintercept = median(fullresults$pgls.r2.rich.sh), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    ## geom_vline(xintercept = median(fullresults$lm.r2.rich.sh), colour = brewer.pal(3, "Set1")[2], linetype = "dashed", size = 1.5) +
    labs(y = element_blank(), x = bquote('R' ^2))+
    facet_wrap(. ~ tree, ncol = 2) +
    theme(axis.text.x = element_text(size = 6)) +
    theme_cowplot()

pvalue.rich.sh.rambarah <-
    ggplot(fullresults) +
    geom_histogram(aes(x = pgls.pvalue.rich.sh, y = ..density..), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 100) +
    geom_histogram(aes(x = lm.pvalue.rich.sh, y = ..density..), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 100) +
    ## geom_vline(xintercept = median(fullresults$pgls.pvalue.rich.sh), colour = brewer.pal(3, "Set1")[1], linetype = "dashed", size = 1.5) +
    labs(y = element_blank(), x = "p-value") +
    facet_wrap(. ~ tree, ncol = 2) +
    theme(axis.text.x = element_text(size = 6)) +
    theme_cowplot()

(scatter.rich.sh.rambarah | (r2.rich.sh.rambarah + pvalue.rich.sh.rambarah)) + plot_annotation(tag_levels = "A")

```


```{r echo = FALSE, fig.cap = "Distribution of p-values of linear (blue bars) and phylogenetic (red bars) ANOVA using different thresholds for classification. A: 50%, B: 60%, C: 80%, D: 100%."}

### Boxplots
## We aggregate mean rates per type for each random dataset

age.data$r.e0 <- geiger::bd.ms(time = age.data$rb.CC.complete.stem, n = age.data$nro_especies, crown = FALSE, epsilon = 0)
age.data$r.e09 <- geiger::bd.ms(time = age.data$rb.CC.complete.stem, n = age.data$nro_especies, crown = FALSE, epsilon = 0.9)

random.data <- lapply(here::here(paste0("output/simulated_datasets/", list.files(here::here("output/simulated_datasets/")))), read.csv, stringsAsFactors = FALSE)

random.types <- data.frame(
    family = age.data$familia,
    r.e0 = age.data$r.e0,
    r.e09 = age.data$r.e09
)

for(i in 1:length(random.data)){
    random.types <- cbind(random.types,
                          random.data[[i]]$type.60
                          )
}

## Removing families that are 100% UNK
random.types <- random.types[-which(random.types[,4] == "UNK"),]

anova.data.r0 <- plyr::ldply(random.types[4:10003], function(x){aggregate(random.types$r.e0, by = list(x), FUN = mean)})
anova.data.r09 <- plyr::ldply(random.types[4:10003], function(x){aggregate(random.types$r.e09, by = list(x), FUN = mean)})


## box.r09 <-
##     ggplot(data = anova.data.r09[is.na(match(anova.data.r09$Group.1, c("OM", "ER"))),]) +
##     geom_boxplot(aes(x = factor(Group.1, levels = c("AM", "EM", "NM", "MIX")), y = x, colour = factor(Group.1, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, outlier.alpha = 1, outlier.shape = 3, size = 1) +
##     geom_point(aes(x = factor(Group.1, levels = c("AM", "EM", "NM", "MIX")), y = x, colour = factor(Group.1, levels = c("AM", "EM", "NM", "MIX"))), alpha = 0.05, position = position_jitterdodge(jitter.width = 2.2, dodge.width = 1)) +
##     stat_summary(aes(x = factor(Group.1, levels = c("AM", "EM", "NM", "MIX")), y = x, colour = factor(Group.1, levels = c("AM", "EM", "NM", "MIX"))), fun.y = mean, geom = "point", colour = "black", size = 2) +
##     theme_cowplot() +
##     theme(legend.position = "none") +
##     scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
##     labs(x = "Mycorrhizal State", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))

## box.r09 <-
##     ggplot(data = family.data.gen[is.na(match(family.data.gen$type.60, c("OM", "ER"))),]) +
##     geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX"))), fill = NA, size = 1.5, outlier.shape = 3, alpha = 0.5) +
##     geom_jitter(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5) +
##     labs(x = "Mycorrhizal State", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"), title = "A") +
##     scale_colour_manual(values = brewer.pal(5, "Set1")[-4]) +
##     theme_cowplot() +
##     theme(legend.position = "none")

## ggsave(filename = "../output/figs/boxplots_anova_myctype_r09_combined.pdf", box.r09, width = 11, height = 5.5, units = "in")

pvalue.phyanova.r09.50 <-
    ggplot(fullresults) +
    geom_histogram(aes(x = phyaov.pvalue.r09.50), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 50) +
    geom_histogram(aes(x = aov.pvalue.r09.50), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 50) +
    geom_vline(xintercept = 0.05, colour = "black", linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value") +
    xlim(-0.005, 1.005) + 
    theme_cowplot()

pvalue.phyanova.r09.60 <-
    ggplot(fullresults) +
    geom_histogram(aes(x = phyaov.pvalue.r09.60), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 50) +
    geom_histogram(aes(x = aov.pvalue.r09.60), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 50) +
    geom_vline(xintercept = 0.05, colour = "black", linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value") +
    xlim(-0.005, 1.005) + 
    theme_cowplot()

pvalue.phyanova.r09.80 <-
    ggplot(fullresults) +
    geom_histogram(aes(x = phyaov.pvalue.r09.80), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 50) +
    geom_histogram(aes(x = aov.pvalue.r09.80), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 50) +
    geom_vline(xintercept = 0.05, colour = "black", linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value") +
    xlim(-0.05, 1.005) + 
    theme_cowplot()

pvalue.phyanova.r09.100 <-
    ggplot(fullresults) +
    geom_histogram(aes(x = phyaov.pvalue.r09.100), fill = brewer.pal(3, "Set1")[1], colour = NA, alpha = 0.5, bins = 50) +
    geom_histogram(aes(x = aov.pvalue.r09.100), fill = brewer.pal(3, "Set1")[2], colour = NA, alpha = 0.5, bins = 50) +
    geom_vline(xintercept = 0.05, colour = "black", linetype = "dashed", size = 1.5) +
    labs(y = "Frequency", x = "p-value") +
    xlim(-0.05, 1.005) + 
    theme_cowplot()

((pvalue.phyanova.r09.50 + pvalue.phyanova.r09.60) / (pvalue.phyanova.r09.80 + pvalue.phyanova.r09.100)) + plot_annotation(tag_levels = "A")

```


# Assigning random families a net diversification of 0 to test for limitations raised by Rabosky \& Benson 2021

## 5% of families

```{r}

### 20% analysis

load(here::here("output/side_analysis_rabosky_benson_5perc.RData"))

side.results <- plyr::ldply(1:length(side.results.5), function(x){side.results.5[[x]][[1]]})

## r_epsilon = 0
hist.pgls.r2.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.r2.r0), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.r2.r0)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.pgls.pvalue.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(pgls.pvalue.r0)), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(pgls.pvalue.r0 + 1e-16))), linetype = "dashed", size = 2, colour = "red") +
    scale_x_continuous(breaks = seq(floor(min(log10(side.results$pgls.pvalue.r0 + 1e-16))), ceiling(max(log10(side.results$pgls.pvalue.r0 + 1e-16))), by = 2)) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()

hist.lm.r2.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.r2.r0), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.r2.r0)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()

hist.lm.pvalue.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(lm.pvalue.r0)), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(lm.pvalue.r0))), linetype = "dashed", size = 2, colour = "blue") +
    #scale_x_continuous(breaks = c(-12, -10, -8, -6, -4), labels = c(bquote('10' ^-12), bquote('10' ^-10), bquote('10' ^-8), bquote('10' ^-6), bquote('10' ^-4))) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()



hist.pgls.slope.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.slope.r0), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.slope.r0)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.lm.slope.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.slope.r0), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.slope.r0)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()


(hist.lm.pvalue.r0 + hist.lm.r2.r0 + hist.lm.slope.r0) / (hist.pgls.pvalue.r0 + hist.pgls.r2.r0 + hist.pgls.slope.r0)

```

```{r}

## r_epslion = 0.9
hist.pgls.r2.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.r2.r09), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.r2.r09)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.pgls.pvalue.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(pgls.pvalue.r09)), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(pgls.pvalue.r09 + 1e-16))), linetype = "dashed", size = 2, colour = "red") +
    #scale_x_continuous(breaks = c(-12.5, -10, -7.5, -5), labels = c(bquote('10' ^-12.5), bquote('10' ^-10), bquote('10' ^-7.5), bquote('10' ^-5))) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()

hist.lm.r2.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.r2.r09), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.r2.r09)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()

hist.lm.pvalue.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(lm.pvalue.r09)), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(lm.pvalue.r09))), linetype = "dashed", size = 2, colour = "blue") +
    #scale_x_continuous(breaks = c(-12, -10, -8, -6, -4), labels = c(bquote('10' ^-12), bquote('10' ^-10), bquote('10' ^-8), bquote('10' ^-6), bquote('10' ^-4))) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()



hist.pgls.slope.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.slope.r09), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.slope.r09)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.lm.slope.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.slope.r09), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.slope.r09)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()


(hist.lm.pvalue.r09 + hist.lm.r2.r09 + hist.lm.slope.r09) / (hist.pgls.pvalue.r09 + hist.pgls.r2.r09 + hist.pgls.slope.r09)

```




## 20% of families

```{r}

### 20% analysis

load(here::here("output/side_analysis_rabosky_benson_20perc.RData"))

side.results <- plyr::ldply(1:length(side.results), function(x){side.results[[x]][[1]]})

## r_epsilon = 0
hist.pgls.r2.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.r2.r0), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.r2.r0)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.pgls.pvalue.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(pgls.pvalue.r0)), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(pgls.pvalue.r0 + 1e-16))), linetype = "dashed", size = 2, colour = "red") +
    scale_x_continuous(breaks = seq(floor(min(log10(side.results$pgls.pvalue.r0 + 1e-16))), ceiling(max(log10(side.results$pgls.pvalue.r0 + 1e-16))), by = 2)) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()

hist.lm.r2.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.r2.r0), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.r2.r0)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()

hist.lm.pvalue.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(lm.pvalue.r0)), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(lm.pvalue.r0))), linetype = "dashed", size = 2, colour = "blue") +
    #scale_x_continuous(breaks = c(-12, -10, -8, -6, -4), labels = c(bquote('10' ^-12), bquote('10' ^-10), bquote('10' ^-8), bquote('10' ^-6), bquote('10' ^-4))) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()



hist.pgls.slope.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.slope.r0), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.slope.r0)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.lm.slope.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.slope.r0), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.slope.r0)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()


(hist.lm.pvalue.r0 + hist.lm.r2.r0 + hist.lm.slope.r0) / (hist.pgls.pvalue.r0 + hist.pgls.r2.r0 + hist.pgls.slope.r0)

```

```{r}

## r_epslion = 0.9
hist.pgls.r2.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.r2.r09), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.r2.r09)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.pgls.pvalue.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(pgls.pvalue.r09)), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(pgls.pvalue.r09 + 1e-16))), linetype = "dashed", size = 2, colour = "red") +
    #scale_x_continuous(breaks = c(-12.5, -10, -7.5, -5), labels = c(bquote('10' ^-12.5), bquote('10' ^-10), bquote('10' ^-7.5), bquote('10' ^-5))) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()

hist.lm.r2.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.r2.r09), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.r2.r09)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()

hist.lm.pvalue.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(lm.pvalue.r09)), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(lm.pvalue.r09))), linetype = "dashed", size = 2, colour = "blue") +
    #scale_x_continuous(breaks = c(-12, -10, -8, -6, -4), labels = c(bquote('10' ^-12), bquote('10' ^-10), bquote('10' ^-8), bquote('10' ^-6), bquote('10' ^-4))) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()



hist.pgls.slope.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.slope.r09), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.slope.r09)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.lm.slope.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.slope.r09), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.slope.r09)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()


(hist.lm.pvalue.r09 + hist.lm.r2.r09 + hist.lm.slope.r09) / (hist.pgls.pvalue.r09 + hist.pgls.r2.r09 + hist.pgls.slope.r09)

```


## 50% of families

```{r}

### 20% analysis

load(here::here("output/side_analysis_rabosky_benson_50perc.RData"))

side.results <- plyr::ldply(1:length(side.results.50), function(x){side.results.50[[x]][[1]]})

## r_epsilon = 0
hist.pgls.r2.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.r2.r0), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.r2.r0)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.pgls.pvalue.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(pgls.pvalue.r0)), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(pgls.pvalue.r0 + 1e-16))), linetype = "dashed", size = 2, colour = "red") +
    scale_x_continuous(breaks = seq(floor(min(log10(side.results$pgls.pvalue.r0 + 1e-16))), ceiling(max(log10(side.results$pgls.pvalue.r0 + 1e-16))), by = 2)) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()

hist.lm.r2.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.r2.r0), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.r2.r0)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()

hist.lm.pvalue.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(lm.pvalue.r0)), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(lm.pvalue.r0))), linetype = "dashed", size = 2, colour = "blue") +
    #scale_x_continuous(breaks = c(-12, -10, -8, -6, -4), labels = c(bquote('10' ^-12), bquote('10' ^-10), bquote('10' ^-8), bquote('10' ^-6), bquote('10' ^-4))) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()



hist.pgls.slope.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.slope.r0), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.slope.r0)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.lm.slope.r0 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.slope.r0), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.slope.r0)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()


(hist.lm.pvalue.r0 + hist.lm.r2.r0 + hist.lm.slope.r0) / (hist.pgls.pvalue.r0 + hist.pgls.r2.r0 + hist.pgls.slope.r0)

```

```{r}

## r_epslion = 0.9
hist.pgls.r2.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.r2.r09), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.r2.r09)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.pgls.pvalue.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(pgls.pvalue.r09)), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(pgls.pvalue.r09 + 1e-16))), linetype = "dashed", size = 2, colour = "red") +
    #scale_x_continuous(breaks = c(-12.5, -10, -7.5, -5), labels = c(bquote('10' ^-12.5), bquote('10' ^-10), bquote('10' ^-7.5), bquote('10' ^-5))) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()

hist.lm.r2.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.r2.r09), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.r2.r09)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()

hist.lm.pvalue.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = log10(lm.pvalue.r09)), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(log10(lm.pvalue.r09))), linetype = "dashed", size = 2, colour = "blue") +
    #scale_x_continuous(breaks = c(-12, -10, -8, -6, -4), labels = c(bquote('10' ^-12), bquote('10' ^-10), bquote('10' ^-8), bquote('10' ^-6), bquote('10' ^-4))) +
    labs(x = "p-value", y = element_blank()) +
    theme_cowplot()



hist.pgls.slope.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = pgls.slope.r09), fill = "red", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(pgls.slope.r09)), linetype = "dashed", size = 2, colour = "red") +
    theme_cowplot()

hist.lm.slope.r09 <-
    ggplot(data = side.results) +
    geom_histogram(aes(x = lm.slope.r09), fill = "blue", colour = NA, alpha = 0.3, bins = 100) +
    geom_vline(aes(xintercept = mean(lm.slope.r09)), linetype = "dashed", size = 2, colour = "blue") +
    theme_cowplot()


(hist.lm.pvalue.r09 + hist.lm.r2.r09 + hist.lm.slope.r09) / (hist.pgls.pvalue.r09 + hist.pgls.r2.r09 + hist.pgls.slope.r09)

```
