---
title: "Supplementary Material - Mujica et al."
output:
  pdf_document:
    toc: false
    fig_caption: true
    latex_engine: lualatex
    number_sections: true
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \floatplacement{table}{H}
  - \floatplacement{verbatim}{H}
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r echo = FALSE}

knitr::opts_chunk$set(echo = FALSE, fig.width = 11, fig.height = 9, warning = FALSE, message = FALSE, fig.pos = 'H')
library("kableExtra")
library("caper")
library("tidyverse")
library("cowplot")
library("RColorBrewer")
rm(list = ls())
load("../output/main_results.RData")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}
options(digits = 3)

```

\beginsupplement

# Analyses with the Species-level dataset

```{r}

## This step is mandatory to avoid problems with overlapping object names
rm(list = ls())
load("../output/suppdata_species_noremarks.RData")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

## Clean dataset - excluding species with any remark

### Boxplots using the clean species-level dataset

```{r fig.height = 11, fig.cap = "Relationship between mycorrhizal type and diversification rates using the thresholds (a) 50%, (b) 60%, (c) 80% and (d) 100% for MIX state assignment using the species-level dataset without remarks. Each panel shows diversification rate estimated with epsilon = 0 (upper) and diversification rate estimated with epsilon = 0.9 (lower). AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates."}

family.data.clean$type.50 <- as.character(family.data.clean$type.50)
family.data.clean$type.60 <- as.character(family.data.clean$type.60)
family.data.clean$type.80 <- as.character(family.data.clean$type.80)
family.data.clean$type.100 <- as.character(family.data.clean$type.100)

box.r0.50 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.50 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))

thresh50 <-
    plot_grid(box.r0.50,
          box.r09.50,
          nrow = 2,
          align = 'v'
          )

box.r0.60 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.60 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh60 <-
    plot_grid(box.r0.60,
          box.r09.60,
          nrow = 2,
          align = 'v'
          )

box.r0.80 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.80 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh80 <-
    plot_grid(box.r0.80,
          box.r09.80,
          nrow = 2,
          align = 'v'
          )

box.r0.100 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e0.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.100 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "ER", "MIX")), y = r.e09.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh100 <-
    plot_grid(box.r0.100,
          box.r09.100,
          nrow = 2,
          align = 'v'
          )

plot_grid(thresh50,
          thresh60,
          thresh80,
          thresh100,
          nrow = 2,
          align = 'hv',
          labels = letters[1:4]
          )

```



### Summary statistics using clean species-level dataset

#### phyANOVA

```{r}

phyaov.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                                   F.r0 = round(c(phyaov.r0.50$F,
                                            phyaov.r0.60$F,
                                            phyaov.r0.80$F,
                                            phyaov.r0.100$F), 3),
                                   pvalue.r0 = round(c(phyaov.r0.50$Pf,
                                                 phyaov.r0.60$Pf,
                                                 phyaov.r0.80$Pf,
                                                 phyaov.r0.100$Pf), 3),
                                   F.r09 = round(c(phyaov.r09.50$F,
                                            phyaov.r09.60$F,
                                            phyaov.r09.80$F,
                                            phyaov.r09.100$F), 3),
                                   pvalue.r09 = round(c(phyaov.r09.50$Pf,
                                                 phyaov.r09.60$Pf,
                                                 phyaov.r09.80$Pf,
                                                 phyaov.r09.100$Pf), 3)
                         )

phyaov.sum[, 2:5] <- apply(phyaov.sum[, 2:5], 2, sig.formatter)

kable_styling(kable(x = phyaov.sum, format = "latex", caption = "summary statistics for phyANOVA for both values of epsilon using the species-level dataset without remarks. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

#### Standard ANOVA

```{r}

std.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                      F.r0 = round(c(summary(aov.r0.50)[[1]]$F[1],
                               summary(aov.r0.60)[[1]]$F[1],
                               summary(aov.r0.80)[[1]]$F[1],
                               summary(aov.r0.100)[[1]]$F[1]), 3),
                      pvalue.r0 = round(c(summary(aov.r0.50)[[1]]$P[1],
                                    summary(aov.r0.60)[[1]]$P[1],
                                    summary(aov.r0.80)[[1]]$P[1],
                                    summary(aov.r0.100)[[1]]$P[1]), 3),
                      F.r09 = round(c(summary(aov.r09.50)[[1]]$F[1],
                                summary(aov.r09.60)[[1]]$F[1],
                                summary(aov.r09.80)[[1]]$F[1],
                                summary(aov.r09.100)[[1]]$F[1]), 3),
                      pvalue.r09 = round(c(summary(aov.r09.50)[[1]]$P[1],
                                     summary(aov.r09.60)[[1]]$P[1],
                                     summary(aov.r09.80)[[1]]$P[1],
                                     summary(aov.r09.100)[[1]]$P[1]), 3)
                      )

std.sum[, 2:5] <- apply(std.sum[, 2:5], 2, sig.formatter)

kable_styling(kable(x = std.sum, format = "latex", caption = "summary statistics for phyANOVA for both values of epsilon using the species-level dataset without remarks. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

### Posthoc tests using clean species-level dataset

#### phyANOVA

```{r}

sum.phyaov <- setNames(
    data.frame(
        threshold = rep(c(50, 60, 80, 100), each = 4),
        Myc.type = rep(c("AM", "EM", "MIX", "NM"), times = 4),
        rbind(phyaov.r0.50$Pt,
              phyaov.r0.60$Pt,
              phyaov.r0.80$Pt,
              phyaov.r0.100$Pt),
        rbind(phyaov.r09.50$Pt,
              phyaov.r09.60$Pt,
              phyaov.r09.80$Pt,
              phyaov.r09.100$Pt)
    ), nm = c("Threshold", "Mycorrhizal.Type", "AM.r0", "EM.r0", "MIX.r0", "NM.r0", "AM.r09", "EM.r09", "MIX.r09", "NM.r09"))
rownames(sum.phyaov) <- NULL

sum.phyaov[, 3:10] <- apply(sum.phyaov[, 3:10], 2, sig.formatter)

kable_styling(kable(x = sum.phyaov, format = "latex", caption = "Pairwise Corrected p-values for phyANOVA using the species-level dataset without remarks. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

#### Standard ANOVA

```{r}

sum.aov <- setNames(
    data.frame(
        pairs = names(TukeyHSD(aov.r0.50)$type.50[,4]),
        round(TukeyHSD(aov.r0.50)$type.50[,4], 3),
        round(TukeyHSD(aov.r0.60)$type.60[,4], 3),
        round(TukeyHSD(aov.r0.80)$type.80[,4], 3),
        round(TukeyHSD(aov.r0.100)$type.100[,4], 3),
        round(TukeyHSD(aov.r09.50)$type.50[,4], 3),
        round(TukeyHSD(aov.r09.60)$type.60[,4], 3),
        round(TukeyHSD(aov.r09.80)$type.80[,4], 3),
        round(TukeyHSD(aov.r09.100)$type.100[,4], 3)
        ), nm = c("Types", "50.r0", "60.r0", "80.r0", "100.r0", "50.r09", "60.r09", "80.r09", "100.r09"))

sum.aov[, 2:9] <- apply(sum.aov[, 2:9], 2, sig.formatter)

kable_styling(kable(x = sum.aov, format = "latex", caption = "Pairwise Corrected p-values for standard ANOVA. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

### Scatterplots

```{r fig.height = 5, fig.cap = "Scatterplots showing the relationship between mycorrhizal diversity index and diversification rates using the species-level dataset without remarks. Diversification rates were estimated with epsilon = 0 (a) and with epsilon = 0.9 (b). The red and blue lines indicate the results of a linear model and a phylogenetic generalized least squares (PGLS) fit, respectively."}

raw.r0 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = shannon, y = r.e0.stem), size = 2) +
    geom_abline(data = data.frame(int = coef(lm.r0)[1], sl = coef(lm.r0)[2], col = "a"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    geom_abline(data = data.frame(int = coef(mod.r0)[1], sl = coef(mod.r0)[2], col = "b"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, linetype = "dashed", show.legend = TRUE) +
    labs(x = "Mycorrhizae Type Shannon Index", y = "Diversification Rate", col = "Model Type") +
    #xlim(0, 1) +
    theme_cowplot() +
    theme(legend.position = "top") +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"))

raw.r09 <-
    ggplot(data = na.omit(family.data.clean)) +
    geom_point(aes(x = shannon, y = r.e09.stem), size = 2) +
    geom_abline(data = data.frame(int = coef(lm.r09)[1], sl = coef(lm.r09)[2], col = "a"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    geom_abline(data = data.frame(int = coef(mod.r09)[1], sl = coef(mod.r09)[2], col = "b"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, linetype = "dashed", show.legend = TRUE) +
    labs(x = "Mycorrhizae Type Shannon Index", y = "Diversification Rate", col = "Model Type") +
    #xlim(0, 1) +
    theme_cowplot() +
    theme(legend.position = "top") +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"))

plot_grid(raw.r0,
          raw.r09,
          ncol = 2,
          align = 'hv',
          axis = 'tlbr',
          labels = letters[1:4]
          )

```

```{r}

reg.sum <- data.frame(epsilon = c(0, 0.9),
                      pvalue.PGLS = c(4.063e-07, 3.483e-07),
                      R2.PGLS = c(0.08904, 0.09007),
                      pvalue.LM = c(1.676e-07, 2.844e-07),
                      R2.LM = c(0.09458, 0.09109))

reg.sum[, 2:5] <- apply(reg.sum[, 2:5], 2, sig.formatter)

kable_styling(kable(x = reg.sum, format = "latex", caption = "Summary statistics for the phylogenetic and parametric regressions using the species-level dataset without remarks. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```


## Full species-level dataset - including species with any remark

```{r}

## This step is mandatory to avoid problems with overlapping object names
rm(list = ls())
load("../output/suppdata_species_withremarks.RData")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

### Boxplots using full species-level dataset

```{r fig.height = 11, fig.cap = "Relationship between mycorrhizal type and diversification rates using the thresholds (a) 50%, (b) 60%, (c) 80% and (d) 100% for MIX state assignment using the species-level dataset with remarks. Each panel shows diversification rate estimated with epsilon = 0 (upper) and diversification rate estimated with epsilon = 0.9 (lower). AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates."}

family.data.clean.rem$type.50 <- as.character(family.data.clean.rem$type.50)
family.data.clean.rem$type.60 <- as.character(family.data.clean.rem$type.60)
family.data.clean.rem$type.80 <- as.character(family.data.clean.rem$type.80)
family.data.clean.rem$type.100 <- as.character(family.data.clean.rem$type.100)

box.r0.rem.50 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.50 != "ER",]) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.rem.50 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.50 != "ER", ]) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))

thresh50 <- 
    plot_grid(box.r0.rem.50,
              box.r09.rem.50,
              nrow = 2,
              align = 'v'
              )

box.r0.rem.60 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.60 != "ER", ]) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.rem.60 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.50 != "ER", ]) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh60 <-
    plot_grid(box.r0.rem.60,
          box.r09.rem.60,
          nrow = 2,
          align = 'v'
          )

box.r0.rem.80 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.80 != "ER", ]) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.rem.80 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.80 != "ER", ]) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))

thresh80 <-
    plot_grid(box.r0.rem.80,
          box.r09.rem.80,
          nrow = 2,
          align = 'v'
          )

box.r0.rem.100 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.100 != "ER", ]) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e0.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.rem.100 <-
    ggplot(data = family.data.clean.rem[family.data.clean.rem$type.100 != "ER", ]) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e09.stem), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))

thresh100 <-
    plot_grid(box.r0.rem.100,
          box.r09.rem.100,
          nrow = 2,
          align = 'v'
          )

plot_grid(thresh50,
          thresh60,
          thresh80,
          thresh100,
          nrow = 2,
          align = 'hv',
          labels = letters[1:4]
          )

```


### Summary statistics using full species-level dataset

#### phyANOVA

```{r}

phyaov.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                         F.r0 = round(c(phyaov.r0.50.rem$F,
                                        phyaov.r0.60.rem$F,
                                        phyaov.r0.80.rem$F,
                                        phyaov.r0.100.rem$F), 3),
                         pvalue.r0 = round(c(phyaov.r0.50.rem$Pf,
                                             phyaov.r0.60.rem$Pf,
                                             phyaov.r0.80.rem$Pf,
                                             phyaov.r0.100.rem$Pf), 3),
                         F.r09 = round(c(phyaov.r09.50.rem$F,
                                         phyaov.r09.60.rem$F,
                                         phyaov.r09.80.rem$F,
                                         phyaov.r09.100.rem$F), 3),
                         pvalue.r09 = round(c(phyaov.r09.50.rem$Pf,
                                              phyaov.r09.60.rem$Pf,
                                              phyaov.r09.80.rem$Pf,
                                              phyaov.r09.100.rem$Pf), 3)
                         )

phyaov.sum[, 2:5] <- apply(phyaov.sum[, 2:5], 2, sig.formatter)

kable_styling(kable(x = phyaov.sum, format = "latex", caption = "summary statistics for phyANOVA for both values of epsilon using the species-level dataset with remarks. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

#### Standard ANOVA

```{r}

std.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                      F.r0 = round(c(summary(aov.r0.50.rem)[[1]]$F[1],
                                     summary(aov.r0.60.rem)[[1]]$F[1],
                                     summary(aov.r0.80.rem)[[1]]$F[1],
                                     summary(aov.r0.100.rem)[[1]]$F[1]), 3),
                      pvalue.r0 = round(c(summary(aov.r0.50.rem)[[1]]$P[1],
                                          summary(aov.r0.60.rem)[[1]]$P[1],
                                          summary(aov.r0.80.rem)[[1]]$P[1],
                                          summary(aov.r0.100.rem)[[1]]$P[1]), 3),
                      F.r09 = round(c(summary(aov.r09.50.rem)[[1]]$F[1],
                                      summary(aov.r09.60.rem)[[1]]$F[1],
                                      summary(aov.r09.80.rem)[[1]]$F[1],
                                      summary(aov.r09.100.rem)[[1]]$F[1]), 3),
                      pvalue.r09 = round(c(summary(aov.r09.50.rem)[[1]]$P[1],
                                           summary(aov.r09.60.rem)[[1]]$P[1],
                                           summary(aov.r09.80.rem)[[1]]$P[1],
                                           summary(aov.r09.100.rem)[[1]]$P[1]), 3)
                      )

std.sum[, 2:5] <- apply(std.sum[, 2:5], 2, sig.formatter)

kable_styling(kable(x = std.sum, format = "latex", caption = "summary statistics for phyANOVA for both values of epsilon using the species-level dataset with remarks. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

### Posthoc tests using full species-level dataset

#### phyANOVA

```{r}

sum.phyaov <- setNames(
    data.frame(
        threshold = rep(c(50, 60, 80, 100), each = 4),
        Myc.type = rep(as.character(c("AM", "EM", "MIX", "NM")), times = 4),
        rbind(phyaov.r0.50.rem$Pt,
              phyaov.r0.60.rem$Pt,
              phyaov.r0.80.rem$Pt,
              phyaov.r0.100.rem$Pt),
        rbind(phyaov.r09.50.rem$Pt,
              phyaov.r09.60.rem$Pt,
              phyaov.r09.80.rem$Pt,
              phyaov.r09.100.rem$Pt),
        stringsAsFactors = FALSE
    ), nm = c("Threshold", "Mycorrhizal.Type", "AM.r0", "EM.r0", "MIX.r0", "NM.r0", "AM.r09", "EM.r09", "MIX.r09", "NM.r09"))
rownames(sum.phyaov) <- NULL

sum.phyaov[, 3:10] <- apply(sum.phyaov[, 3:10], 2, sig.formatter)

kable_styling(kable(x = sum.phyaov, format = "latex", caption = "Pairwise Corrected p-values for phyANOVA using the species-level dataset with remarks. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

#### Standard ANOVA

```{r}

sum.aov <- setNames(
    data.frame(
        pairs = names(TukeyHSD(aov.r0.50.rem)$type.50[,4]),
        round(TukeyHSD(aov.r0.50.rem)$type.50[,4], 3),
        round(TukeyHSD(aov.r0.60.rem)$type.60[,4], 3),
        round(TukeyHSD(aov.r0.80.rem)$type.80[,4], 3),
        round(TukeyHSD(aov.r0.100.rem)$type.100[,4], 3),
        round(TukeyHSD(aov.r09.50.rem)$type.50[,4], 3),
        round(TukeyHSD(aov.r09.60.rem)$type.60[,4], 3),
        round(TukeyHSD(aov.r09.80.rem)$type.80[,4], 3),
        round(TukeyHSD(aov.r09.100.rem)$type.100[,4], 3)
        ), nm = c("Types", "50.r0", "60.r0", "80.r0", "100.r0", "50.r09", "60.r09", "80.r09", "100.r09"))

sum.aov[, 2:9] <- apply(sum.aov[, 2:9], 2, sig.formatter)

kable_styling(kable(x = sum.aov, format = "latex", caption = "Pairwise Corrected p-values for standard ANOVA. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

### Scatterplots

```{r fig.height = 5, fig.cap = "Scatterplots showing the relationship between mycorrhizal diversity index and diversification rates using the species-level dataset with remarks. Diversification rates were estimated with epsilon = 0 (a) and with epsilon = 0.9 (b). The red and blue lines indicate the results of a linear model and a phylogenetic generalized least squares (PGLS) fit, respectively."}

raw.r0.rem <-
    ggplot(data = na.omit(family.data.clean.rem)) +
    geom_point(aes(x = shannon, y = r.e0.stem), size = 2) +
    geom_abline(data = data.frame(int = coef(lm.r0.rem)[1], sl = coef(lm.r0.rem)[2], col = "a"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    geom_abline(data = data.frame(int = coef(mod.r0.rem)[1], sl = coef(mod.r0.rem)[2], col = "b"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, linetype = "dashed", show.legend = TRUE) +
    labs(x = "Mycorrhizae Type Shannon Index", y = "Diversification Rate", col = "Model Type") +
    #xlim(0, 1) +
    theme_cowplot() +
    theme(legend.position = "top") +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"))

raw.r09.rem <-
    ggplot(data = na.omit(family.data.clean.rem)) +
    geom_point(aes(x = shannon, y = r.e09.stem), size = 2) +
    geom_abline(data = data.frame(int = coef(lm.r09.rem)[1], sl = coef(lm.r09.rem)[2], col = "a"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    geom_abline(data = data.frame(int = coef(mod.r09.rem)[1], sl = coef(mod.r09.rem)[2], col = "b"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, linetype = "dashed", show.legend = TRUE) +
    labs(x = "Mycorrhizae Type Shannon Index", y = "Diversification Rate", col = "Model Type") +
    #xlim(0, 1) +
    theme_cowplot() +
    theme(legend.position = "top") +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"))

plot_grid(raw.r0.rem,
          raw.r09.rem,
          ncol = 2,
          align = 'hv',
          axis = 'tlbr',
          labels = letters[1:4]
          )

```

```{r}

reg.sum <- data.frame(epsilon = c(0, 0.9),
                      pvalue.PGLS = c(5.265e-05, 4.013e-05),
                      R2.PGLS = c(0.06444, 0.06653),
                      pvalue.LM = c(2.528e-07, 4.977e-07),
                      R2.LM = c(0.09354, 0.08898))

reg.sum[, 2:5] <- apply(reg.sum[, 2:5], 2, sig.formatter)

kable_styling(kable(x = reg.sum, format = "latex", caption = "Summary statistics for the phylogenetic and parametric regressions using the species-level dataset with remarks. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

# Adding randomly 20% of misassignment of mycorrhizal type

```{r}

rm(list = ls())
load("../output/results_20perc.RData")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

## Regressions

```{r fig.cap = "Distribution of (a) p-values and (b) R² values for the same analyses of the main paper, but using all 50 replicate datasets with mycorrhizal types for 20% of species randomly sampled. Dashed vertical lines represent the corresponding empirical value from the main analysis."}

main.data.reg <- data.frame(
    epsilon = c(0, 0.9, 0, 0.9),
    model = c("PGLS", "PGLS", "LM", "LM"),
    pvalue = c(2.336e-12, 7.124e-13, 5.908e-15, 5.464e-15),
    R2 = c(0.1251, 0.1307, 0.1524, 0.1527)
)

pvalue.20perc.reg <-
    ggplot(data = pvalue.reg) +
    geom_histogram(aes(x = log10(pvalue), y = ..density.., fill = model), alpha = 0.3, colour = "white") +
    geom_density(aes(x = log10(pvalue), colour = model, fill = model), alpha = 0.3) +
    geom_vline(data = main.data.reg, aes(xintercept = log10(pvalue), group = model), colour = "black", linetype = "dashed") +
    facet_grid(model ~ epsilon) +
    scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set1")[2:1]) +
    scale_fill_manual(values = RColorBrewer::brewer.pal(3, "Set1")[2:1]) +
    labs(x = "p-value (log10)", y = "Density") +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10))

pvalue.20perc.r2 <-
    ggplot(data = pvalue.reg) +
    geom_histogram(aes(x = R2, y = ..density.., fill = model), alpha = 0.3, colour = "white") +
    geom_density(aes(x = R2, fill = model, colour = model), alpha = 0.3) +
    geom_vline(data = main.data.reg, aes(xintercept = R2, group = model), colour = "black", linetype = "dashed") +
    facet_grid(model ~ epsilon) +
    scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set1")[2:1]) +
    scale_fill_manual(values = RColorBrewer::brewer.pal(3, "Set1")[2:1]) +
    labs(x = "R²", y = "Density") +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10))

plot_grid(pvalue.20perc.reg,
          pvalue.20perc.r2,
          ncol = 2,
          align = 'h',
          labels = letters[1:2]
          )

```



## ANOVAs

```{r fig.height = 11, fig.cap = "Distribution of p-values of (a) epsilon = 0 and (b) epsilon = 0.9 for the both phylogenetic and standard ANOVA, but using all 50 replicate datasets with mycorrhizal types for 20% of species randomly sampled. Dashed vertical lines represent the corresponding empirical value from the main analysis."}

main.data.aov <- data.frame(
    threshold = rep(c(50, 60, 80, 100), 2),
    model = rep(c("phy", "std"), each = 4),
    pvalue.r0 = c(0.089, 0.013, 0.001, 0.001),
    pvalue.r09 = c(0.162, 0.007, 0.001, 0.001)
    )

pvalue.20perc.aov.r0 <-
    ggplot(data = pvalue.aov) +
    geom_histogram(aes(x = pvalue.r0, y = ..density.., fill = model), alpha = 0.3, colour = "white") +
    geom_vline(data = main.data.aov, aes(xintercept = pvalue.r0, group = model), colour = "black", linetype = "dashed") +
    facet_grid(threshold ~ model) +
    scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set1")[2:1]) +
    scale_fill_manual(values = RColorBrewer::brewer.pal(3, "Set1")[2:1]) +
    labs(x = "p-value", y = "Density") +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10))

pvalue.20perc.aov.r09 <-
    ggplot(data = pvalue.aov) +
    geom_histogram(aes(x = pvalue.r09, y = ..density.., fill = model), alpha = 0.3, colour = "white") +
    geom_vline(data = main.data.aov, aes(xintercept = pvalue.r09, group = model), colour = "black", linetype = "dashed") +
    facet_grid(threshold ~ model) +
    scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Set1")[2:1]) +
    scale_fill_manual(values = RColorBrewer::brewer.pal(3, "Set1")[2:1]) +
    labs(x = "p-value", y = "Density") +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10))

plot_grid(pvalue.20perc.aov.r0,
          pvalue.20perc.aov.r09,
          ncol = 2,
          align = 'h',
          labels = letters[1:2]
          )

```

# Analyses with the genus-level dataset

```{r echo = FALSE}

rm(list = ls())
load("../output/main_results.RData")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

## Analysis per genera excluding 100% MIX families

### Boxplots using clean genus-level dataset

```{r fig.height = 11, fig.cap = "Relationship between mycorrhizal type and diversification rates using the genus-level dataset excluding 100% MIX families. a) diversification rate estimated with epsilon = 0 and b) diversification rate estimated with epsilon = 0.9. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates."}

family.data.gen$type.50 <- as.character(family.data.gen$type.50)
family.data.gen$type.60 <- as.character(family.data.gen$type.60)
family.data.gen$type.80 <- as.character(family.data.gen$type.80)
family.data.gen$type.100 <- as.character(family.data.gen$type.100)

family.data.gen <- family.data.gen[-which(is.na(family.data.gen$r.e0)),]

box.r0.50 <-
    ggplot(data = family.data.gen[family.data.gen$type.50 != "ER", ]) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e0, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e0), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.50 <-
    ggplot(data = family.data.gen[family.data.gen$type.50 != "ER", ]) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e09), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh50 <-
    plot_grid(box.r0.50,
          box.r09.50,
          nrow = 2,
          align = 'v'
          )

box.r0.60 <-
    ggplot(data = family.data.gen[family.data.gen$type.60 != "ER", ]) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e0, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e0), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.60 <-
    ggplot(data = family.data.gen[family.data.gen$type.60 != "ER", ]) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh60 <-
    plot_grid(box.r0.60,
          box.r09.60,
          nrow = 2,
          align = 'v'
          )

box.r0.80 <-
    ggplot(data = family.data.gen[family.data.gen$type.80 != "ER", ]) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e0, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e0), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.80 <-
    ggplot(data = family.data.gen[family.data.gen$type.80 != "ER", ]) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e09), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh80 <-
    plot_grid(box.r0.80,
          box.r09.80,
          nrow = 2,
          align = 'v'
          )

box.r0.100 <-
    ggplot(data = family.data.gen[family.data.gen$type.100 != "ER", ]) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e0, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e0), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.100 <-
    ggplot(data = family.data.gen[family.data.gen$type.100 != "ER", ]) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e09), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
    scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh100 <-
    plot_grid(box.r0.100,
          box.r09.100,
          nrow = 2,
          align = 'v'
          )

plot_grid(thresh50,
          thresh60,
          thresh80,
          thresh100,
          nrow = 2,
          align = 'hv',
          labels = letters[1:4]
          )

```

```{r}

fam.class <- data.frame(Threshold = c("50", "60", "80", "100"),
                        rbind(table(family.data.gen$type.50),
                              table(family.data.gen$type.60),
                              table(family.data.gen$type.80),
                              table(family.data.gen$type.100)))

# fam.class <- apply(fam.class, 2, sig.formatter)

kable_styling(kable(x = fam.class, format = "latex", caption = "Mycorrhizal type assigned to each family based on 4 different percentage thresholds (50, 60, 80 and 100) using the genus-level dataset excluding 100% MIX families. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```




### Summary statistics using clean genus-level dataset

#### phyANOVA

```{r}

phyaov.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                                   F.r0 = round(c(phyaov.r0.50$F,
                                            phyaov.r0.60$F,
                                            phyaov.r0.80$F,
                                            phyaov.r0.100$F), 3),
                                   pvalue.r0 = round(c(phyaov.r0.50$Pf,
                                                 phyaov.r0.60$Pf,
                                                 phyaov.r0.80$Pf,
                                                 phyaov.r0.100$Pf), 3),
                                   F.r09 = round(c(phyaov.r09.50$F,
                                            phyaov.r09.60$F,
                                            phyaov.r09.80$F,
                                            phyaov.r09.100$F), 3),
                                   pvalue.r09 = round(c(phyaov.r09.50$Pf,
                                                 phyaov.r09.60$Pf,
                                                 phyaov.r09.80$Pf,
                                                 phyaov.r09.100$Pf), 3)
                         )

phyaov.sum[, 2:5] <- apply(phyaov.sum[, 2:5], 2, sig.formatter)

kable_styling(kable(x = phyaov.sum, format = "latex", caption = "Summary statistics for phyANOVA for both values of epsilon using the genus-level dataset excluding 100% MIX families. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

#### Standard ANOVA

```{r}

aov.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                      F.r0 = round(c(summary(aov.r0.50)[[1]]$F[1],
                                     summary(aov.r0.60)[[1]]$F[1],
                                     summary(aov.r0.80)[[1]]$F[1],
                                     summary(aov.r0.100)[[1]]$F[1]), 3),
                      pvalue.r0 = round(c(summary(aov.r0.50)[[1]]$P[1],
                                          summary(aov.r0.60)[[1]]$P[1],
                                          summary(aov.r0.80)[[1]]$P[1],
                                          summary(aov.r0.100)[[1]]$P[1]), 3),
                      F.r09 = round(c(summary(aov.r09.50)[[1]]$F[1],
                                      summary(aov.r09.60)[[1]]$F[1],
                                      summary(aov.r09.80)[[1]]$F[1],
                                      summary(aov.r09.100)[[1]]$F[1]), 3),
                      pvalue.r09 = round(c(summary(aov.r09.50)[[1]]$P[1],
                                           summary(aov.r09.60)[[1]]$P[1],
                                           summary(aov.r09.80)[[1]]$P[1],
                                           summary(aov.r09.100)[[1]]$P[1]), 3)
                      )

aov.sum[, 2:5] <- apply(aov.sum[, 2:5], 2, sig.formatter)

kable_styling(kable(x = aov.sum, format = "latex", caption = "summary statistics for standard ANOVA for both values of epsilon using the genus-level dataset excluding 100% MIX families. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```
### Posthoc tests

#### phyANOVA

```{r}

sum.phyaov <- setNames(
                        data.frame(
                            threshold = rep(c(50, 60, 80, 100), each = 4),
                            Myc.type = rep(c("AM", "EM", "MIX", "NM"), times = 4),
                        rbind(phyaov.r0.50$Pt,
                              phyaov.r0.60$Pt,
                              phyaov.r0.80$Pt,
                              phyaov.r0.100$Pt),
                        rbind(phyaov.r09.50$Pt,
                              phyaov.r09.60$Pt,
                              phyaov.r09.80$Pt,
                              phyaov.r09.100$Pt)
                        ), nm = c("Threshold", "Mycorrhizal.Type", "AM.r0", "EM.r0", "MIX.r0", "NM.r0", "AM.r09", "EM.r09", "MIX.r09", "NM.r09"))
rownames(sum.phyaov) <- NULL

sum.phyaov[, 3:10] <- apply(sum.phyaov[, 3:10], 2, sig.formatter)

kable_styling(kable(x = sum.phyaov, format = "latex", caption = "Pairwise Corrected p-values for phyANOVA using the genus-level dataset excluding 100% MIX families. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

#### Standard ANOVA

```{r}

sum.aov <- setNames(
    data.frame(
        pairs = names(TukeyHSD(aov.r0.50)$type.50[,4]),
        round(TukeyHSD(aov.r0.50)$type.50[,4], 3),
        round(TukeyHSD(aov.r0.60)$type.60[,4], 3),
        round(TukeyHSD(aov.r0.80)$type.80[,4], 3),
        round(TukeyHSD(aov.r0.100)$type.100[,4], 3),
        round(TukeyHSD(aov.r09.50)$type.50[,4], 3),
        round(TukeyHSD(aov.r09.60)$type.60[,4], 3),
        round(TukeyHSD(aov.r09.80)$type.80[,4], 3),
        round(TukeyHSD(aov.r09.100)$type.100[,4], 3)
        ), nm = c("Types", "50.r0", "60.r0", "80.r0", "100.r0", "50.r09", "60.r09", "80.r09", "100.r09"))

sum.aov[, 2:9] <- apply(sum.aov[, 2:9], 2, sig.formatter)

kable_styling(kable(x = sum.aov, format = "latex", caption = "Pairwise Corrected p-values for standard ANOVA using the genus-level dataset excluding 100% MIX families. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

## Genus-level full dataset - including all families

```{r}

## This step is mandatory to avoid problems with overlapping object names
rm(list = ls())
load("../output/suppdata_genus_full.RData")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

### Boxplots using genus-level full dataset

```{r fig.height = 11, fig.cap = "Relationship between mycorrhizal type and diversification rates using the thresholds (a) 50%, (b) 60%, (c) 80% and (d) 100% for MIX state assignment using the genus-level dataset including all families. Each panel shows diversification rate estimated with epsilon = 0 (upper) and diversification rate estimated with epsilon = 0.9 (lower). AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates."}

family.data.gen$type.50 <- as.character(family.data.gen$type.50)
family.data.gen$type.60 <- as.character(family.data.gen$type.60)
family.data.gen$type.80 <- as.character(family.data.gen$type.80)
family.data.gen$type.100 <- as.character(family.data.gen$type.100)

family.data.gen <- family.data.gen[-which(is.na(family.data.gen$r.e0)),]

box.r0.50 <-
    ggplot(data = family.data.gen[family.data.gen$type.50 != "ER",]) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e0, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e0), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.50 <-
    ggplot(data = family.data.gen[family.data.gen$type.50 != "ER",]) +
    geom_point(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.50, levels = c("AM", "EM", "NM", "MIX")), y = r.e09), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh50 <- 
    plot_grid(box.r0.50,
              box.r09.50,
              nrow = 2,
              align = 'v'
              )

box.r0.60 <-
    ggplot(data = family.data.gen[family.data.gen$type.60 != "ER",]) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e0, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e0), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.60 <-
    ggplot(data = family.data.gen[family.data.gen$type.50 != "ER",]) +
    geom_point(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = r.e09), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh60 <-
    plot_grid(box.r0.60,
          box.r09.60,
          nrow = 2,
          align = 'v'
          )

box.r0.80 <-
    ggplot(data = family.data.gen[family.data.gen$type.80 != "ER",]) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e0, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e0), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.80 <-
    ggplot(data = family.data.gen[family.data.gen$type.80 != "ER",]) +
    geom_point(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.80, levels = c("AM", "EM", "NM", "MIX")), y = r.e09), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh80 <-
    plot_grid(box.r0.80,
          box.r09.80,
          nrow = 2,
          align = 'v'
          )

box.r0.100 <-
    ggplot(data = family.data.gen[family.data.gen$type.100 != "ER",]) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e0, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e0), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0 )"))

box.r09.100 <-
    ggplot(data = family.data.gen[family.data.gen$type.100 != "ER",]) +
    geom_point(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e09, colour = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.75, position = position_jitterdodge(jitter.width = 0.75, dodge.width = 1)) +
    geom_boxplot(aes(x = factor(type.100, levels = c("AM", "EM", "NM", "MIX")), y = r.e09), fill = NA, colour = "darkgrey", outlier.alpha = 1) +
    theme_cowplot() +
    theme(legend.position = "none", axis.title.y = element_text(size = 10)) +
scale_colour_manual(values = brewer.pal(5, "Set1")[c(1:3, 5)]) +
    labs(x = "Mycorrhizal Type", y = expression("Diversification Rate ("~epsilon~" = 0.9 )"))


thresh100 <-
    plot_grid(box.r0.100,
          box.r09.100,
          nrow = 2,
          align = 'v'
          )

plot_grid(thresh50,
          thresh60,
          thresh80,
          thresh100,
          nrow = 2,
          align = 'hv',
          labels = letters[1:4]
          )

```



### Summary statistics

#### phyANOVA

```{r}

phyaov.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                                   F.r0 = round(c(phyaov.r0.50$F,
                                            phyaov.r0.60$F,
                                            phyaov.r0.80$F,
                                            phyaov.r0.100$F), 3),
                                   pvalue.r0 = round(c(phyaov.r0.50$Pf,
                                                 phyaov.r0.60$Pf,
                                                 phyaov.r0.80$Pf,
                                                 phyaov.r0.100$Pf), 3),
                                   F.r09 = round(c(phyaov.r09.50$F,
                                            phyaov.r09.60$F,
                                            phyaov.r09.80$F,
                                            phyaov.r09.100$F), 3),
                                   pvalue.r09 = round(c(phyaov.r09.50$Pf,
                                                 phyaov.r09.60$Pf,
                                                 phyaov.r09.80$Pf,
                                                 phyaov.r09.100$Pf), 3)
                         )

phyaov.sum <- apply(phyaov.sum, 2, sig.formatter)

kable_styling(kable(x = phyaov.sum, format = "latex", caption = "summary statistics for phyANOVA for both values of epsilon using the genus-level dataset including all families. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

#### Standard ANOVA

```{r}

aov.sum <- data.frame(Threshold = c(50, 60, 80, 100),
                      F.r0 = round(c(summary(aov.r0.50)[[1]]$F[1],
                                     summary(aov.r0.60)[[1]]$F[1],
                                     summary(aov.r0.80)[[1]]$F[1],
                                     summary(aov.r0.100)[[1]]$F[1]), 3),
                      pvalue.r0 = round(c(summary(aov.r0.50)[[1]]$P[1],
                                          summary(aov.r0.60)[[1]]$P[1],
                                          summary(aov.r0.80)[[1]]$P[1],
                                          summary(aov.r0.100)[[1]]$P[1]), 3),
                      F.r09 = round(c(summary(aov.r09.50)[[1]]$F[1],
                                      summary(aov.r09.60)[[1]]$F[1],
                                      summary(aov.r09.80)[[1]]$F[1],
                                      summary(aov.r09.100)[[1]]$F[1]), 3),
                      pvalue.r09 = round(c(summary(aov.r09.50)[[1]]$P[1],
                                           summary(aov.r09.60)[[1]]$P[1],
                                           summary(aov.r09.80)[[1]]$P[1],
                                           summary(aov.r09.100)[[1]]$P[1]), 3)
                      )
aov.sum <- apply(aov.sum, 2, sig.formatter)

kable_styling(kable(x = aov.sum, format = "latex", caption = "summary statistics for standard ANOVA for both values of epsilon using the genus-level dataset including all families. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

### Posthoc tests using genus-level full dataset

#### phyANOVA

```{r}

sum.phyaov <- setNames(
                        data.frame(
                            threshold = rep(c(50, 60, 80, 100), each = 4),
                            Myc.type = rep(c("AM", "EM", "MIX", "NM"), times = 4),
                        rbind(phyaov.r0.50$Pt,
                              phyaov.r0.60$Pt,
                              phyaov.r0.80$Pt,
                              phyaov.r0.100$Pt),
                        rbind(phyaov.r09.50$Pt,
                              phyaov.r09.60$Pt,
                              phyaov.r09.80$Pt,
                              phyaov.r09.100$Pt)
                        ), nm = c("Threshold", "Mycorrhizal.Type", "AM.r0", "EM.r0", "MIX.r0", "NM.r0", "AM.r09", "EM.r09", "MIX.r09", "NM.r09"))
rownames(sum.phyaov) <- NULL

sum.phyaov[, 3:10] <- apply(sum.phyaov[, 3:10], 2, sig.formatter)

kable_styling(kable(x = sum.phyaov, format = "latex", caption = "Pairwise Corrected p-values for phyANOVA using the genus-level dataset including all families. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

#### Standard ANOVA

```{r}

sum.aov <- setNames(
    data.frame(
        pairs = names(TukeyHSD(aov.r0.50)$type.50[,4]),
        round(TukeyHSD(aov.r0.50)$type.50[,4], 3),
        round(TukeyHSD(aov.r0.60)$type.60[,4], 3),
        round(TukeyHSD(aov.r0.80)$type.80[,4], 3),
        round(TukeyHSD(aov.r0.100)$type.100[,4], 3),
        round(TukeyHSD(aov.r09.50)$type.50[,4], 3),
        round(TukeyHSD(aov.r09.60)$type.60[,4], 3),
        round(TukeyHSD(aov.r09.80)$type.80[,4], 3),
        round(TukeyHSD(aov.r09.100)$type.100[,4], 3)
        ), nm = c("Pairs", "50.r0", "60.r0", "80.r0", "100.r0", "50.r09", "60.r09", "80.r09", "100.r09"))

sum.aov[, 2:9] <- apply(sum.aov[, 2:9], 2, sig.formatter)

kable_styling(kable(x = sum.aov, format = "latex", caption = "Pairwise Corrected p-values for standard ANOVA. Significant values are highlighted in bold.", row.names = FALSE, escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

### Scatterplots using genus-level full dataset

```{r fig.height = 5, fig.cap = "Scatterplots showing the relationship between mycorrhizal diversity index and diversification rates using the genus-level dataset including all families. Diversification rates were estimated with epsilon = 0 (a) and with epsilon = 0.9 (b). The red and blue lines indicate the results of a linear model and a phylogenetic generalized least squares (PGLS) fit, respectively."}

## Scatter - MTDI vs. Net Diversification
mtdi.r0 <-
    ggplot(data = na.omit(family.data.gen)) +
    geom_point(aes(x = shannon, y = r.e0), size = 2) +
    geom_abline(data = data.frame(int = coef(lm.r0)[1], sl = coef(lm.r0)[2], col = "a"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, linetype = "dashed", show.legend = TRUE) +
    geom_abline(data = data.frame(int = coef(mod.r0)[1], sl = coef(mod.r0)[2], col = "b"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    labs(x = "Mycorrhizal Type Shannon Index", y = "Diversification Rate", col = "Model Type") +
    #xlim(0, 1) +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"), direction = -1) +
    theme_cowplot() +
    theme(legend.position = "top")

mtdi.r09 <-
    ggplot(data = na.omit(family.data.gen)) +
    geom_point(aes(x = shannon, y = r.e0), size = 2) +
    geom_abline(data = data.frame(int = coef(lm.r09)[1], sl = coef(lm.r09)[2], col = "a"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, linetype = "dashed", show.legend = TRUE) +
    geom_abline(data = data.frame(int = coef(mod.r09)[1], sl = coef(mod.r09)[2], col = "b"), mapping = aes(intercept = int, slope = sl, colour = col), size = 1.5, show.legend = TRUE) +
    labs(x = "Mycorrhizal Type Shannon Index", y = "Diversification Rate", col = "Model Type") +
    #xlim(0, 1) +
    scale_colour_brewer(palette = "Set1", labels = c("Linear Model", "PGLS"), direction = -1) +
    theme_cowplot() +
    theme(legend.position = "top")

plot_grid(mtdi.r0,
          mtdi.r09,
          ncol = 2,
          align = 'hv',
          axis = 'tlbr',
          labels = letters[1:2]
          )

```

```{r}

reg.sum <- data.frame(epsilon = c(0, 0.9),
                      pvalue.PGLS = c(4.637e-08, 6.858e-08),
                      R2.PGLS = c(0.07343, 0.07157),
                      pvalue.LM = c(6.325e-10, 4.512e-09),
                      R2.LM = c(0.09319, 0.08402))

reg.sum[, 2:5] <- apply(reg.sum[, 2:5], 2, sig.formatter)

kable_styling(kable(x = reg.sum, format = "latex", caption = "Summary statistics for the phylogenetic and parametric regressions using the genus-level dataset including all families. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

# Phylogenetic signal of diversification rates, age, and richness

```{r}

rm(list = ls())
load("../output/main_results.RData")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

```{r}

data.phylo.d50 <- family.data.gen[, c("family", "type.50")]
data.phylo.d60 <- family.data.gen[, c("family", "type.60")]
data.phylo.d80 <- family.data.gen[, c("family", "type.80")]
data.phylo.d100 <- family.data.gen[, c("family", "type.100")]

## Threshold 50%

data.phylo.d50$dummy.am <- ifelse(data.phylo.d50$type.50 == "AM", 1, 0)
data.phylo.d50$dummy.em <- ifelse(data.phylo.d50$type.50 == "EM", 1, 0)
data.phylo.d50$dummy.mix <- ifelse(data.phylo.d50$type.50 == "MIX", 1, 0)
data.phylo.d50$dummy.nm <- ifelse(data.phylo.d50$type.50 == "NM", 1, 0)

d.am.50 <- phylo.d(data.phylo.d50, data.pgls$phy, names.col = family, binvar = dummy.am)
d.em.50 <- phylo.d(data.phylo.d50, data.pgls$phy, names.col = family, binvar = dummy.em)
d.mix.50 <- phylo.d(data.phylo.d50, data.pgls$phy, names.col = family, binvar = dummy.mix)
d.nm.50 <- phylo.d(data.phylo.d50, data.pgls$phy, names.col = family, binvar = dummy.nm)

## Threshold 60%

data.phylo.d60$dummy.am <- ifelse(data.phylo.d60$type.60 == "AM", 1, 0)
data.phylo.d60$dummy.em <- ifelse(data.phylo.d60$type.60 == "EM", 1, 0)
data.phylo.d60$dummy.mix <- ifelse(data.phylo.d60$type.60 == "MIX", 1, 0)
data.phylo.d60$dummy.nm <- ifelse(data.phylo.d60$type.60 == "NM", 1, 0)

d.am.60 <- phylo.d(data.phylo.d60, data.pgls$phy, names.col = family, binvar = dummy.am)
d.em.60 <- phylo.d(data.phylo.d60, data.pgls$phy, names.col = family, binvar = dummy.em)
d.mix.60 <- phylo.d(data.phylo.d60, data.pgls$phy, names.col = family, binvar = dummy.mix)
d.nm.60 <- phylo.d(data.phylo.d60, data.pgls$phy, names.col = family, binvar = dummy.nm)

## Threshold 80%

data.phylo.d80$dummy.am <- ifelse(data.phylo.d80$type.80 == "AM", 1, 0)
data.phylo.d80$dummy.em <- ifelse(data.phylo.d80$type.80 == "EM", 1, 0)
data.phylo.d80$dummy.mix <- ifelse(data.phylo.d80$type.80 == "MIX", 1, 0)
data.phylo.d80$dummy.nm <- ifelse(data.phylo.d80$type.80 == "NM", 1, 0)

d.am.80 <- phylo.d(data.phylo.d80, data.pgls$phy, names.col = family, binvar = dummy.am)
d.em.80 <- phylo.d(data.phylo.d80, data.pgls$phy, names.col = family, binvar = dummy.em)
d.mix.80 <- phylo.d(data.phylo.d80, data.pgls$phy, names.col = family, binvar = dummy.mix)
d.nm.80 <- phylo.d(data.phylo.d80, data.pgls$phy, names.col = family, binvar = dummy.nm)

## Threshold 100%

data.phylo.d100$dummy.am <- ifelse(data.phylo.d100$type.100 == "AM", 1, 0)
data.phylo.d100$dummy.em <- ifelse(data.phylo.d100$type.100 == "EM", 1, 0)
data.phylo.d100$dummy.mix <- ifelse(data.phylo.d100$type.100 == "MIX", 1, 0)
data.phylo.d100$dummy.nm <- ifelse(data.phylo.d100$type.100 == "NM", 1, 0)

d.am.100 <- phylo.d(data.phylo.d100, data.pgls$phy, names.col = family, binvar = dummy.am)
d.em.100 <- phylo.d(data.phylo.d100, data.pgls$phy, names.col = family, binvar = dummy.em)
d.mix.100 <- phylo.d(data.phylo.d100, data.pgls$phy, names.col = family, binvar = dummy.mix)
d.nm.100 <- phylo.d(data.phylo.d100, data.pgls$phy, names.col = family, binvar = dummy.nm)

phylod.table <- data.frame(
                            Threshold = rep(c(50, 60, 80, 100), each = 2),
                            model = rep(c("random", "BM"), times = 4),
                            AM = c(unlist(d.am.50[c("Pval1", "Pval0")]),
                                   unlist(d.am.60[c("Pval1", "Pval0")]),
                                   unlist(d.am.80[c("Pval1", "Pval0")]),
                                   unlist(d.am.100[c("Pval1", "Pval0")])),
                            EM = c(unlist(d.em.50[c("Pval1", "Pval0")]),
                                   unlist(d.em.60[c("Pval1", "Pval0")]),
                                   unlist(d.em.80[c("Pval1", "Pval0")]),
                                   unlist(d.em.100[c("Pval1", "Pval0")])),
                            MIX = c(unlist(d.mix.50[c("Pval1", "Pval0")]),
                                   unlist(d.mix.60[c("Pval1", "Pval0")]),
                                   unlist(d.mix.80[c("Pval1", "Pval0")]),
                                   unlist(d.mix.100[c("Pval1", "Pval0")])),
                            NM = c(unlist(d.nm.50[c("Pval1", "Pval0")]),
                                   unlist(d.nm.60[c("Pval1", "Pval0")]),
                                   unlist(d.nm.80[c("Pval1", "Pval0")]),
                                   unlist(d.nm.100[c("Pval1", "Pval0")]))
)

phylod.table[, 3:6] <- apply(phylod.table[, 3:6], 2, sig.formatter)

kable_styling(kable(x = phylod.table, format = "latex", caption = "P-values for test of phylogenetic signal (D) of each mycorrhizal type using the genus-level dataset including all families. Significant values are highlighted in bold.", escape = FALSE, digits = 3), latex_options = "HOLD_position")

```

```{r}

kable_styling(kable(x = data.frame(Variable = c("r.e0", "r.e09", "Stem.Age", "Richness"),
                                   Lambda = c(paste0("\\textbf{", round(summary(phylosig.r0)$param[2], 3), "}"),
                                              paste0("\\textbf{", round(summary(phylosig.r09)$param[2], 3), "}"), 
                                              paste0("\\textbf{", summary(phylosig.age)$param[2], "}"),
                                              summary(phylosig.rich)$param[2]
                                              )),
                    format = "latex", caption = "Phylogenetic signal of the four response variables using the genus-level dataset including all families. Significant values are highlighted in bold.", escape = FALSE), latex_options = "HOLD_position"
              )

```

# Replicating main analyses with tree from Harris and Davies (2016)

```{r}

rm(list = ls())
load("../output/harris_davies_results.RData")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

```{r fig.height = 7, fig.cap = "Association between placement of families in both Zanne's (left) and Harris and Davies' (right) phylogenies, indicating differences in the topologies."}

cophyloplot(fulltree, tree, assoc = cophylo.myco$assoc, use.edge.length = TRUE, show.tip.label = FALSE, space = 500)

```

```{r fig.height = 5, fig.cap = "Relationship between net diversification calculated in our study and with the rates from Harris and Davies (2016) The rates were estimated by the authors using the same source for species richness as the main analysis, and the stem ages were obtained from their phylogeny. a: epsilon = 0; b: epsilon = 0.9. Red lines: LM fit; blue dashed lines: identity fit."}

r0.corr <-
    ggplot(data.muj) +
    geom_point(aes(x = new.phylo.r.e0, y = r.e0)) +
    geom_smooth(aes(x = new.phylo.r.e0, y = r.e0), method = "lm", se = FALSE, colour = "red") +
    geom_abline(slope = 1, intercept = 0, colour = "blue", linetype = "dashed") +
    labs(x = "Harris and Davies\n(epsilon = 0)", y = "Zanne et al\n(epsilon = 0)") +
    cowplot::theme_cowplot()

r09.corr <-
    ggplot(data.muj) +
    geom_point(aes(x = new.phylo.r.e09, y = r.e09)) +
    geom_smooth(aes(x = new.phylo.r.e09, y = r.e09), method = "lm", se = FALSE, colour = "red") +
    geom_abline(slope = 1, intercept = 0, colour = "blue", linetype = "dashed") +
    labs(x = "Harris and Davies\n(epsilon = 0.9)", y = "Zanne et al\n(epsilon = 0.9)") +
    cowplot::theme_cowplot()

plot_grid(r0.corr,
          r09.corr,
          ncol = 2,
          align = 'h',
          labels = letters[1:2]
          )

```

```{r fig.cap = "Main analyses replicated using Harris and Davies (2016) phylogeny. The rates were estimated by the authors using the same source for species richness as the main analysis, and the stem ages were obtained from their phylogeny. Panels a and c: epsilon = 0; panels b and d: epsilon = 0.9. Red lines: PGLS fit; blue dashed lines: LM fit."}

sh.r0.np <-
    ggplot(data = data.muj.plot) +
    geom_point(aes(x = shannon, y = new.phylo.r.e0)) +
    geom_abline(intercept = coef(lm.np.r0)[1], slope = coef(lm.np.r0)[2], colour = brewer.pal(3, "Set1")[2], size = 2, linetype = "dashed") +
    geom_abline(intercept = coef(mod.np.r0)[1], slope = coef(mod.np.r0)[2], colour = brewer.pal(3, "Set1")[1], size = 2) +
    labs(x = "Mycorrhizal Type Diversity Index", y = "NetDiversification\n(epsilon = 0)", colour = "Model") +
    theme_cowplot() +
    theme(legend.position = "top")

sh.r09.np <-
    ggplot(data = data.muj.plot) +
    geom_point(aes(x = shannon, y = new.phylo.r.e0)) +
    geom_abline(intercept = coef(lm.np.r09)[1], slope = coef(lm.np.r09)[2], colour = brewer.pal(3, "Set1")[2], size = 2, linetype = "dashed") +
    geom_abline(intercept = coef(mod.np.r09)[1], slope = coef(mod.np.r09)[2], colour = brewer.pal(3, "Set1")[1], size = 2) +
    labs(x = "Mycorrhizal Type Diversity Index", y = "NetDiversification\n(epsilon = 0.9)", colour = "Model") +
    theme_cowplot() +
    theme(legend.position = "top")


aov.r0.np <-
    ggplot(data = data.aov) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = new.phylo.r.e0), colour = "lightgrey") +
    geom_jitter(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = new.phylo.r.e0, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5) +
    labs(x = "Mycorrhizal Type", y = "NetDiversification\n(epsilon = 0)") +
    scale_colour_manual(values = RColorBrewer::brewer.pal(5, "Set1")[c(1:3, 5)]) +
    theme_cowplot() +
    theme(legend.position = "none")


aov.r09.np <-
    ggplot(data = data.aov) +
    geom_boxplot(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = new.phylo.r.e09), colour = "lightgrey") +
    geom_jitter(aes(x = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), y = new.phylo.r.e09, colour = factor(type.60, levels = c("AM", "EM", "NM", "MIX")), size = shannon), alpha = 0.5) +
    labs(x = "Mycorrhizal Type", y = "NetDiversification\n(epsilon = 0.9)") +
    scale_colour_manual(values = RColorBrewer::brewer.pal(5, "Set1")[c(1:3, 5)]) +
    theme_cowplot() +
    theme(legend.position = "none")

plot_grid(sh.r0.np,
          sh.r09.np,
          aov.r0.np,
          aov.r09.np,
          ncol = 2,
          align = 'hv',
          labels = letters[1:4]
          )

```

```{r}

np.sum <-
    data.frame(epsilon = c(0, 0.9),
                     pvalue.PGLS = c("\\textbf{2.724e-13}", "\\textbf{3.581e-13}"),
                     R2.PGLS = c(0.1349, 0.1336),
                     pvalue.LM = c("\\textbf{1.345e-14}", "\\textbf{1.228e-14}"),
                     R2.LM = c(0.1464, 0.1468),
                     pvalue.stdaov = c("\\textbf{0.019}", "\\textbf{0.02}"),
                     pvalue.phyaov = c("\\textbf{0.0005}", "\\textbf{0.0003}")
                     )

kable_styling(kable(x = np.sum,
                    format = "latex", caption = "Summary statistics for linear models and ANOVA using Harris and Davies (2016) phylogeny.", escape = FALSE), latex_options = "HOLD_position"
              )

```
