---
title: "Supplementary Material"
output:
  pdf_document:
    toc: true
    fig_caption: true
    latex_engine: lualatex
    number_sections: false
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \floatplacement{table}{H}
  - \floatplacement{verbatim}{H}
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  - \usepackage{lscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
---

```{r echo = FALSE}

knitr::opts_chunk$set(echo = FALSE, fig.width = 11, fig.height = 9, warning = FALSE, message = FALSE, fig.pos = 'H')
library("kableExtra")
library("caper")
library("tidyverse")
library("cowplot")
library("RColorBrewer")
library("ape")
library("RColorBrewer")
library("geiger")
library("phytools")
library("ggrepel")
library("patchwork")
library("plyr")
library("here")
here::i_am("manuscript/supp_mat_2023.Rmd")
rm(list = ls())
load(here::here("output/main_results.RData"))
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}
options(digits = 3)

```

\beginsupplement

\blandscape


# Appendix S3 - Randomly adding 20% of misassignment of mycorrhizal type into the data obtained from genus-level list, with rates calculated using Zanne et al. 2014

```{r fig.height = 4, fig.cap = "Distribution of R\\textsuperscript{2}, p-value and slope value distributions for $\\epsilon$ = 0 after adding 20% misassignment of mycorrhizal type."}

### 20% analysis
rm(list = ls())
load(here::here("output/hist_20perc.RData"))

(hist.pgls.pvalue.r0 + hist.pgls.r2.r0 + hist.pgls.slope.r0) +
    plot_annotation(tag_levels = "A")

```

```{r fig.height = 4, fig.cap = "Distribution of R\\textsuperscript{2}, p-value and slope value distributions for $\\epsilon$ = 0.9 after adding 20% misassignment of mycorrhizal type."}

(hist.pgls.pvalue.r09 + hist.pgls.r2.r09 + hist.pgls.slope.r09) +
    plot_annotation(tag_levels = "A")

```

# Appendix S5 - Analyses with the genus-level dataset

## Thresholds for mycorrhizal state assignment

```{r}

## This step is mandatory to avoid problems with overlapping object names
rm(list = ls())
load(here::here("output/anova_boxplots.RData"))
##source("../R/anova_plots.R")
sig.formatter <- function(x){cell_spec(x, format = "latex", bold = ifelse(x < 0.05, TRUE, FALSE))}

```

\blandscape

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 50\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0. Panel (A): Boxplot with a randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value for phylogenetic ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r0.50) + (((pvalue.phyanova.r0.50) / posthoc.r0.50) +
               plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 60\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value for phylogenetic ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r0.60) + (((pvalue.phyanova.r0.60) / posthoc.r0.60) +
               plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 80\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value for phylogenetic ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r0.80) + (((pvalue.phyanova.r0.80) / posthoc.r0.80) +
               plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 100\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value for phylogenetic ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r0.100) + (((pvalue.phyanova.r0.100) / posthoc.r0.100) +
                plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 50\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0.9. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value for phylogenetic ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r09.50) + (((pvalue.phyanova.r09.50) / posthoc.r09.50) +
                plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 60\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0.9. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value for phylogenetic ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05. This is the same as figure 2 of the main text and was left there only for easy comparison."}

(box.r09.60) + (((pvalue.phyanova.r09.60) / posthoc.r09.60) +
                plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 80\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0.9. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value for phylogenetic ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r09.80) + (((pvalue.phyanova.r09.80) / posthoc.r09.80) +
                plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

```{r fig.width = 11, fig.height = 7, fig.cap = "Relationship between mycorrhizal type and diversification rates using the threshold 100\\% for MIX state assignment using the genus-level dataset and with $\\epsilon$ = 0.9. Panel (A): Boxplot of randomly chosen dataset replicate showing the distribution of diversification rates for each mycorrhizal state. AM: Arbuscular mycorrhiza, EM: Ectomycorrhiza, NM: non-mycorrhizal and MIX (families with no dominance of any specific mycorrhizal association). The size of the points indicates the Mycorrhizal Type Diversity Index value for each lineage, indicating a predominance of larger indices with higher diversification rates. Panel (B): Histogram of p-value for phylogenetic ANOVA for all 10000 datasets. Panel (C): p-value of post-hoc tests of all pairwise combinations of mycorrhizal states using all 10000 datasets. In panels (B) and (C) the vertical dashed line correspond to 0.05"}

(box.r09.100) + (((pvalue.phyanova.r09.100) / posthoc.r09.100) +
                 plot_layout(heights = c(3, 7))) +
    plot_layout(widths = c(6, 4))

```

\elandscape


### Summary statistics

```{r}

phyaov.sum <- data.frame(Quantile = c("Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"),
                         as.vector(summary(fullresults$phyaov.pvalue.r0.50)),
                         as.vector(summary(fullresults$phyaov.pvalue.r0.60)),
                         as.vector(summary(fullresults$phyaov.pvalue.r0.80)),
                         as.vector(summary(fullresults$phyaov.pvalue.r0.100)),
                         as.vector(summary(fullresults$phyaov.pvalue.r09.50)),
                         as.vector(summary(fullresults$phyaov.pvalue.r09.60)),
                         as.vector(summary(fullresults$phyaov.pvalue.r09.80)),
                         as.vector(summary(fullresults$phyaov.pvalue.r09.100))
                         )

names(phyaov.sum)[2:9] <- c("50\\%", "60\\%", "80\\%", "100\\%", "50\\%", "60\\%", "80\\%", "100\\%")
phyaov.sum[, 2:9] <- apply(phyaov.sum[, 2:9], 2, sig.formatter)
#names(phyaov.sum)[2:5] <- c("F (e = 0)", "p-value (e = 0)", "F (e = 0.9)", "p-value (e = 0.9)")

kable(x = phyaov.sum, format = "latex", caption = "Summary statistics for phyANOVA for both values of $\\epsilon$ using the genus-level dataset including all families to test for differences in diversification rates. Significant values are highlighted in bold.", escape = FALSE, digits = 3, col.names = c("Quantile", "50\\%", "60\\%", "80\\%", "100\\%", "50\\%", "60\\%", "80\\%", "100\\%"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
  add_header_above(c(" ", "e = 0" = 4, "e = 0.9" = 4))

```

### Posthoc tests using genus-level full dataset

\newpage

```{r}

sum.phyaov <- setNames(
    data.frame(
        epsilon = rep(c(0, 0.9), each = 24),
        threshold = rep(rep(c("50\\%", "60\\%", "80\\%", "100\\%"), each = 6), 2),
        state1 <- c("AM", "AM", "EM", "AM", "EM", "MIX"),
        state2 <- c("EM", "MIX", "MIX", "NM", "NM", "NM"),
        plyr::ldply(list(phyaov.r0.50, phyaov.r0.60, phyaov.r0.80, phyaov.r0.100, phyaov.r09.50, phyaov.r09.60, phyaov.r09.80, phyaov.r09.100), function(x){aggregate(na.omit(x)$value, by = list(na.omit(x)$variable, na.omit(x)$type.2), FUN = summary)$x})),
        nm = c("Epsilon", "Threshold", "State 1", "State 2", "Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"))

rownames(sum.phyaov) <- NULL

#sum.phyaov[, 5:10] <- apply(sum.phyaov[, 3:10], 2, sig.formatter)
#names(sum.phyaov)[3:10] <- c("AM (e = 0)", "EM (e = 0)", "MIX (e = 0)", "NM (e = 0)", "AM (e = 0.9)", "EM (e = 0.9)", "MIX (e = 0.9)", "NM (e = 0.9)")

kable(x = sum.phyaov, format = "latex", caption = "Summary statistics for pairwise corrected p-value distributions for phyANOVA using all 10000 genus-level dataset. Values were rounded to the third decimal place.", row.names = FALSE, escape = FALSE, digits = 3, col.names = c("Epsilon", "Threshold", "State 1", "State 2", "Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F)

```

```{r fig.height = 5, fig.cap = "Scatterplots showing a randomly selected dataset replicate (out of 10000) to show the relationship between mycorrhizal state diversity index and diversification rates estimated with $\\epsilon$ (relative extinction fraction) = 0.9. The red lines indicate the phylogenetic generalized least squares (PGLS) fit for 1000 datasets of seed plant mycorrhizal states.  Panel b show the frequency distribution of R2 and p-values of PGLS, dashed lines show the median for the distribution."}

## Scatter - MSDI vs. Net Diversification
load(here::here("output/scatter_r0.RData"))
(scatter.mtdi.r0 | ((r2.pgls.r0 + pvalue.pgls.r0))) +
    plot_annotation(tag_levels = "A")

```


```{r}

reg.sum <- data.frame(Quantile = c("Min", "1st Quart", "Median", "Mean", "3rd Quart", "Max"),
                      pvalue.r0.phy = as.vector(summary(fullresults$pgls.pvalue.r0)),
                      R2.r0.phy = as.vector(summary(fullresults$pgls.r2.r0)),
                      pvalue.r09.phy = as.vector(summary(fullresults$pgls.pvalue.r09)),
                      R2.r09.phy = as.vector(summary(fullresults$pgls.r2.r09))## ,
                      ## pvalue.r0.lm = as.vector(summary(fullresults$lm.pvalue.r0)),
                      ## R2.r0.lm = as.vector(summary(fullresults$lm.r2.r0)),
                      ## pvalue.r09.lm = as.vector(summary(fullresults$lm.pvalue.r09)),
                      ## R2.r09.lm = as.vector(summary(fullresults$lm.r2.r09))
                      )

#reg.sum[, 2:5] <- apply(reg.sum[, 2:5], 2, sig.formatter)

kable(x = reg.sum, format = "latex", caption = "Summary statistics for p-value and R\\textsuperscript{2} for the phylogenetic regression using all 10000 genus-level datasets.", escape = FALSE, digits = 3, col.names = c("Quantile", "p-value", "R\\textsuperscript{2}", "p-value", "R\\textsuperscript{2}"), align = "c") %>%
  kable_styling(#latex_options = "HOLD_position",
                full_width = F) %>%
    add_header_above(c(" ", "e = 0" = 2, "e = 0.9" = 2))##  %>%
    ## add_header_above(c(" ", "PGLS" = 4, "LM" = 4))

```

# Appendix S8 - Percentage of MIX species

```{r fig.width = 11, fig.height = 7, fig.cap = "Distribution of percentage of species in each family that are classified as MIX. Zoom panel indicates the same distribution but excluding families without any MIX species."}

library("ggforce")
mix.sp <- read.csv(here::here("data/family_data_genus_final.csv"), stringsAsFactors = FALSE)
mix.sp$zoom.bin <- "a"
mix.sp$zoom.bin[mix.sp$MIX.raw.perc != 0] <- "b"

fullhist <-
    ggplot(mix.sp) +
    geom_histogram(aes(x = MIX.raw.perc, fill = zoom.bin, colour = zoom.bin), bins = nrow(mix.sp), size = 2) +
    labs(x = "Proportion of MIX species", y = "Frequency", title = "A") +
    scale_colour_manual(values = c("a" = "black", "b" = "red")) +
    scale_fill_manual(values = c("a" = "black", "b" = "red")) +
    theme_cowplot() +
    theme(legend.position = "none")

mix.hist <-
    ggplot(mix.sp[mix.sp$MIX.raw.perc != 0,]) +
    geom_histogram(aes(x = MIX.raw.perc, fill = zoom.bin, colour = zoom.bin), bins = nrow(mix.sp[mix.sp$MIX.raw.perc != 0,]), colour = "red", fill = "red") +
    labs(x = "Proportion of MIX species", y = "Frequency", title = "B") +
    theme_cowplot()

fullhist + annotation_custom(grob = ggplotGrob(mix.hist), xmin = 0.3, xmax = 1, ymin = 100, ymax = 300)

## ggsave(filename = "./output/figs/hist_mix_proportion.pdf")

```


# Appendix S12 - Analyses using rates from Igea & Tanentzap 2020

<!-- ## Lambda per type - species level -->

```{r lambda-per-type-sp, echo = FALSE, fig.height = 6, fig.cap = "Distribution of lambda values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."}

library("caper")
fulltree <- read.tree(here::here("data/fam_tree_family_full.tre"))
fulltree$node.label <- NULL
fulltree <- drop.tip(fulltree, "Didiereaceae")

bamm.rates <- read.csv(here::here("output/bamm_rates.csv"))
family.lambda <- aggregate(bamm.rates$lambda, by = list(bamm.rates$family), FUN = "mean", na.rm = TRUE)
family.mu <- aggregate(bamm.rates$mu, by = list(bamm.rates$family), FUN = "mean", na.rm = TRUE)

family.data <- read.csv(here::here("output/full_family_data_2021.csv"))

family.data$lambda <- family.lambda$x[match(family.data$family, family.lambda$Group.1)]
family.data$mu <- family.mu$x[match(family.data$family, family.lambda$Group.1)]

family.data <- family.data[-match(c("Orchidaceae", "Ericaceae"), family.data$family),]

species.data <- read.table(here::here("output/musse/species_data/moldata.txt"), na.strings = "NA", sep = "\t", header = TRUE)
genus.data <- read.csv(here::here("data/genus_database_Soudzilovskaia.csv"), na.strings = "NA")

sp.data <- bamm.rates[bamm.rates$species %in% species.data$species, ]
sp.data$type <- species.data$trait.num[match(sp.data$species, species.data$species)]

tree.pruned <- drop.tip(fulltree, tip = fulltree$tip.label[is.na(match(fulltree$tip.label, family.data$family))])
data.pgls <- comparative.data(tree.pruned, family.data, names.col = "family")

## ggplot(sp.data) +
##     geom_boxplot(aes(x = type, y = lambda, colour = factor(type)), fill = NA) +
##     geom_jitter(aes(x = type, y = lambda, colour = factor(type)), alpha = 0.1) +
##     stat_summary(aes(x = type, y = lambda), fun = median, geom = "point", size = 4) +
##     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) +
##     theme_cowplot() +
##     theme(legend.position = "top")

```

<!-- ## Mu per type - species level -->

<!-- ```{r mu-per-type-sp, echo = FALSE, fig.height = 6, fig.cap = "Distribution of mu values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."} -->

<!-- ggplot(sp.data) + -->
<!--     geom_boxplot(aes(x = type, y = mu, colour = factor(type)), fill = NA) + -->
<!--     geom_jitter(aes(x = type, y = mu, colour = factor(type)), alpha = 0.1) + -->
<!--     stat_summary(aes(x = type, y = mu), fun = median, geom = "point", size = 4) + -->
<!--     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) + -->
<!--     theme_cowplot() + -->
<!--     theme(legend.position = "top") -->

<!-- ``` -->


## Net Diversification per type - species level

```{r netdiv-per-type-sp, echo = FALSE, fig.height = 6, fig.cap = "Distribution of netdiv values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."}

ggplot(sp.data) +
    geom_boxplot(aes(x = type, y = lambda - mu, colour = factor(type)), fill = NA) +
    geom_jitter(aes(x = type, y = lambda - mu, colour = factor(type)), alpha = 0.1) +
    stat_summary(aes(x = type, y = lambda - mu), fun = median, geom = "point", size = 4) +
    scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) +
    theme_cowplot() +
    theme(legend.position = "top")

```

<!-- # Lambda per type - family level -->

<!-- ## 50% Threshold -->

<!-- ```{r lambda-per-type-fam-50, echo = FALSE, fig.height = 6, fig.cap = "Distribution of lambda values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."} -->

<!-- ggplot(family.data) + -->
<!--     geom_boxplot(aes(x = type.50, y = lambda, colour = factor(type.50)), fill = NA) + -->
<!--     geom_jitter(aes(x = type.50, y = lambda, colour = factor(type.50)), alpha = 0.1) + -->
<!--     stat_summary(aes(x = type.50, y = lambda), fun = median, geom = "point", size = 4) + -->
<!--     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) + -->
<!--     theme_cowplot() + -->
<!--     theme(legend.position = "top") -->

<!-- ``` -->

<!-- ## 60% Threshold -->

<!-- ```{r lambda-per-type-fam-60, echo = FALSE, fig.height = 6, fig.cap = "Distribution of lambda values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."} -->

<!-- ggplot(family.data) + -->
<!--     geom_boxplot(aes(x = type.60, y = lambda, colour = factor(type.60)), fill = NA) + -->
<!--     geom_jitter(aes(x = type.60, y = lambda, colour = factor(type.60)), alpha = 0.1) + -->
<!--     stat_summary(aes(x = type.60, y = lambda), fun = median, geom = "point", size = 4) + -->
<!--     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) + -->
<!--     theme_cowplot() + -->
<!--     theme(legend.position = "top") -->

<!-- ``` -->

<!-- ## 80% Threshold -->

<!-- ```{r lambda-per-type-fam-80, echo = FALSE, fig.height = 6, fig.cap = "Distribution of lambda values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."} -->

<!-- ggplot(family.data) + -->
<!--     geom_boxplot(aes(x = type.80, y = lambda, colour = factor(type.80)), fill = NA) + -->
<!--     geom_jitter(aes(x = type.80, y = lambda, colour = factor(type.80)), alpha = 0.1) + -->
<!--     stat_summary(aes(x = type.80, y = lambda), fun = median, geom = "point", size = 4) + -->
<!--     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) + -->
<!--     theme_cowplot() + -->
<!--     theme(legend.position = "top") -->

<!-- ``` -->

<!-- ## 100% Threshold -->

<!-- ```{r lambda-per-type-fam-100, echo = FALSE, fig.height = 6, fig.cap = "Distribution of lambda values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."} -->

<!-- ggplot(family.data) + -->
<!--     geom_boxplot(aes(x = type.100, y = lambda, colour = factor(type.100)), fill = NA) + -->
<!--     geom_jitter(aes(x = type.100, y = lambda, colour = factor(type.100)), alpha = 0.1) + -->
<!--     stat_summary(aes(x = type.100, y = lambda), fun = median, geom = "point", size = 4) + -->
<!--     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) + -->
<!--     theme_cowplot() + -->
<!--     theme(legend.position = "top") -->

<!-- ``` -->


<!-- # Mu per type - family level -->

<!-- ## 50% Threshold -->

<!-- ```{r mu-per-type-fam-50, echo = FALSE, fig.height = 6, fig.cap = "Distribution of mu values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."} -->

<!-- ggplot(family.data) + -->
<!--     geom_boxplot(aes(x = type.50, y = mu, colour = factor(type.50)), fill = NA) + -->
<!--     geom_jitter(aes(x = type.50, y = mu, colour = factor(type.50)), alpha = 0.1) + -->
<!--     stat_summary(aes(x = type.50, y = mu), fun = median, geom = "point", size = 4) + -->
<!--     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) + -->
<!--     theme_cowplot() + -->
<!--     theme(legend.position = "top") -->

<!-- ``` -->

<!-- ## 60% Threshold -->

<!-- ```{r mu-per-type-fam-60, echo = FALSE, fig.height = 6, fig.cap = "Distribution of mu values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."} -->

<!-- ggplot(family.data) + -->
<!--     geom_boxplot(aes(x = type.60, y = mu, colour = factor(type.60)), fill = NA) + -->
<!--     geom_jitter(aes(x = type.60, y = mu, colour = factor(type.60)), alpha = 0.1) + -->
<!--     stat_summary(aes(x = type.60, y = mu), fun = median, geom = "point", size = 4) + -->
<!--     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) + -->
<!--     theme_cowplot() + -->
<!--     theme(legend.position = "top") -->

<!-- ``` -->

<!-- ## 80% Threshold -->

<!-- ```{r mu-per-type-fam-80, echo = FALSE, fig.height = 6, fig.cap = "Distribution of mu values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."} -->

<!-- ggplot(family.data) + -->
<!--     geom_boxplot(aes(x = type.80, y = mu, colour = factor(type.80)), fill = NA) + -->
<!--     geom_jitter(aes(x = type.80, y = mu, colour = factor(type.80)), alpha = 0.1) + -->
<!--     stat_summary(aes(x = type.80, y = mu), fun = median, geom = "point", size = 4) + -->
<!--     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) + -->
<!--     theme_cowplot() + -->
<!--     theme(legend.position = "top") -->

<!-- ``` -->

<!-- ## 100% Threshold -->

<!-- ```{r mu-per-type-fam-100, echo = FALSE, fig.height = 6, fig.cap = "Distribution of mu values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."} -->

<!-- ggplot(family.data) + -->
<!--     geom_boxplot(aes(x = type.100, y = mu, colour = factor(type.100)), fill = NA) + -->
<!--     geom_jitter(aes(x = type.100, y = mu, colour = factor(type.100)), alpha = 0.1) + -->
<!--     stat_summary(aes(x = type.100, y = mu), fun = median, geom = "point", size = 4) + -->
<!--     scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) + -->
<!--     theme_cowplot() + -->
<!--     theme(legend.position = "top") -->

<!-- ``` -->



## Net Diversification per type - family level

### 50% Threshold

```{r netdiv-per-type-sp-50, echo = FALSE, fig.height = 6, fig.cap = "Distribution of netdiv values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."}

ggplot(family.data) +
    geom_boxplot(aes(x = type.50, y = lambda - mu, colour = factor(type.50)), fill = NA) +
    geom_jitter(aes(x = type.50, y = lambda - mu, colour = factor(type.50)), alpha = 0.1) +
    stat_summary(aes(x = type.50, y = lambda - mu), fun = median, geom = "point", size = 4) +
    scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) +
    theme_cowplot() +
    theme(legend.position = "top")

```

### phylANOVA - 50% Threshold

```{r phyanova-bamm-nd-50}

phyaov.nd.50 <- caper::pgls((lambda - mu) ~ type.50, data = data.pgls, lambda = "ML")
summary(phyaov.nd.50)

```


### 60% Threshold

```{r netdiv-per-type-sp-60, echo = FALSE, fig.height = 6, fig.cap = "Distribution of netdiv values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."}

ggplot(family.data) +
    geom_boxplot(aes(x = type.60, y = lambda - mu, colour = factor(type.60)), fill = NA) +
    geom_jitter(aes(x = type.60, y = lambda - mu, colour = factor(type.60)), alpha = 0.1) +
    stat_summary(aes(x = type.60, y = lambda - mu), fun = median, geom = "point", size = 4) +
    scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) +
    theme_cowplot() +
    theme(legend.position = "top")

```

### phylANOVA - 60% Threshold

```{r phyanova-bamm-nd-60}

phyaov.nd.60 <- caper::pgls((lambda - mu) ~ type.60, data = data.pgls, lambda = "ML")
summary(phyaov.nd.60)

```

### 80% Threshold

```{r netdiv-per-type-sp-80, echo = FALSE, fig.height = 6, fig.cap = "Distribution of netdiv values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."}

ggplot(family.data) +
    geom_boxplot(aes(x = type.80, y = lambda - mu, colour = factor(type.80)), fill = NA) +
    geom_jitter(aes(x = type.80, y = lambda - mu, colour = factor(type.80)), alpha = 0.1) +
    stat_summary(aes(x = type.80, y = lambda - mu), fun = median, geom = "point", size = 4) +
    scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) +
    theme_cowplot() +
    theme(legend.position = "top")

```

### phylANOVA - 80% Threshold

```{r phyanova-bamm-nd-80}

phyaov.nd.80 <- caper::pgls((lambda - mu) ~ type.80, data = data.pgls, lambda = "ML")
summary(phyaov.nd.80)

```

### 100% Threshold

```{r netdiv-per-type-sp-100, echo = FALSE, fig.height = 6, fig.cap = "Distribution of netdiv values per type of mycorrhiza. Black dots indicate the median net diversification rate for each type."}

ggplot(family.data) +
    geom_boxplot(aes(x = type.100, y = lambda - mu, colour = factor(type.100)), fill = NA) +
    geom_jitter(aes(x = type.100, y = lambda - mu, colour = factor(type.100)), alpha = 0.1) +
    stat_summary(aes(x = type.100, y = lambda - mu), fun = median, geom = "point", size = 4) +
    scale_colour_brewer(palette = "Set1", labels = c("AM", "EM", "ER", "MIX", "NM", "OM")) +
    theme_cowplot() +
    theme(legend.position = "top")

```

### phylANOVA - 100% Threshold

```{r phyanova-bamm-nd-100}

phyaov.nd.100 <- caper::pgls((lambda - mu) ~ type.100, data = data.pgls, lambda = "ML")
summary(phyaov.nd.100)

```

\newpage

## PGLS shannon vs. rates

```{r pgls-bamm-rates, echo = FALSE}

library("caper")
fulltree <- read.tree(here::here("data/fam_tree_family_full.tre"))
fulltree$node.label <- NULL
fulltree <- drop.tip(fulltree, "Didiereaceae")

tree.pruned <- drop.tip(fulltree, tip = fulltree$tip.label[is.na(match(fulltree$tip.label, family.data$family))])
data.pgls <- comparative.data(tree.pruned, family.data, names.col = "family")

mod.nd <- caper::pgls((lambda - mu) ~ shannon, data = data.pgls, lambda = "ML")
summary(mod.nd)

mod.lambda <- caper::pgls(lambda ~ shannon, data = data.pgls, lambda = "ML")
summary(mod.lambda)

mod.mu <- caper::pgls(mu ~ shannon, data = data.pgls, lambda = "ML")
summary(mod.mu)

```

## PGLS plot shannon vs. net diversification

```{r pgls-plot, fig.height = 5, fig.cap = "Relationship between MSDI and net diversification rates estimated using BAMM."}

ggplot(family.data) +
    geom_point(aes(x = shannon, y = lambda - mu), alpha = 0.3) +
    geom_abline(intercept = coef(summary(mod.nd))[1,1], slope = coef(summary(mod.nd))[2,1], colour = "red", size = 1.5) +
    theme_cowplot()

```

# Appendix S13 - Randomly adding 20% of misassignment of mycorrhizal type into the data obtained from genus-level list, using rates from Igea & Tanentzap (2020)

```{r fig.height = 4, fig.cap = "Distribution of R\\textsuperscript{2}, p-value and slope value distributions for $\\lambda$ after adding 20% misassignment of mycorrhizal type."}

### 20% analysis
rm(list = ls())
load(here::here("output/hist_20perc_igeatanen_2023.RData"))

(hist.pgls.pvalue.lambda + hist.pgls.r2.lambda + hist.pgls.slope.lambda) +
    plot_annotation(tag_levels = "A")

```

```{r fig.height = 4, fig.cap = "Distribution of R\\textsuperscript{2}, p-value and slope value distributions for $\\mu$ after adding 20% misassignment of mycorrhizal type."}

### 20% analysis
(hist.pgls.pvalue.mu + hist.pgls.r2.mu + hist.pgls.slope.mu) +
    plot_annotation(tag_levels = "A")

```

```{r fig.height = 4, fig.cap = "Distribution of R\\textsuperscript{2}, p-value and slope value distributions for $r$ after adding 20% misassignment of mycorrhizal type."}

### 20% analysis
(hist.pgls.pvalue.netdiv + hist.pgls.r2.netdiv + hist.pgls.slope.netdiv) +
    plot_annotation(tag_levels = "A")

```

